---
title: "Street Trees Independent Variables"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: html_document
---
This file contains independent variables for street tree data analysis 
Uses BNIA neighborhood delineations or Community Statistical Area (CSA)

# 0 set up
```{r include=FALSE}
# load packages

library(tidyverse)
library(sf) #simple features - vector GIS support
library(mapview) #quick web maps 
library(tidylog) #makes dplyr and tidyr verbose

#working directory
getwd()
list.files()
list.files('../BaltimoreStreetTreeProject_Large_Data/')
list.files('../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)

#`%nin%` <- Negate(in)

#test for push notifications
```

Read in BNIA csv files with demographic data

#1 BNIA CSA race demographics ## use 2020
```{r include=FALSE}

# PercentAsian <-  
#   read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Asian_(Non-Hispanic).csv') %>%   select(CSA2020, pasi20) %>% 
#   rename(PercAsian = pasi20)
# 
# PercentLatino <- 
#   read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Hispanic.csv') %>% 
#   select(CSA2020, phisp20) %>% 
#   rename(PercLatino = phisp20)


PercentBlk <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Black_African-American_(Non-Hispanic).csv') %>% 
  select(CSA2020, paa20) %>% 
  rename(PercBlk = paa20)

ggplot(data = PercentBlk, aes(x = PercBlk)) + geom_histogram()

PercentWhite <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_White_Caucasian_(Non-Hispanic).csv') %>% 
  select(CSA2020, pwhite20) %>% 
  rename(PercWhite = pwhite20) 

ggplot(data = PercentWhite, aes(x = PercWhite)) + geom_histogram()
```

#2 Median income
```{r}
Income <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Median_Household_Income.csv') %>% 
  select(CSA2020, mhhi20) 

ggplot(data = Income, aes(x = mhhi20)) + geom_histogram()
```

#3 Education - percent of population 25 and over with a Bachelor's Degree or above
```{r}
Edu <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Percent_Population_(25_Years_and_over)_with_a_Bachelor_s_Degree_or_Above.csv') %>% 
  select(CSA2020, bahigher20) 

```

#4 Temperature
```{r}
#read in temperature by neighborhood csv 
(temperature_by_neigh <- 
    read_csv('input_data/NB_af_temp_2024-05-07.csv' 
             , col_names = c('CSA2020', 'avg_temp', 'n_pixel'))
)

NBtemperature <- temperature_by_neigh %>% 
  select(-n_pixel)
```

#5 % Vacant and abandoned residential properties
```{r}
PercVacant <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Percentage_of_Residential_Properties_that_are_Vacant_and_Abandoned.csv') %>% 
  select(CSA2020, vacant20)
```

#6 Land Cover BNIA total NB area
```{r}
LandCover <- read_csv('../BaltimoreStreetTreeProject_Large_Data\\neighs_bnia.csv'
           , col_names = c('CSA2020'
                           , 'n'
                           , 'lc_class')) %>%
  filter(lc_class != 0) %>% 
  filter(!is.na(lc_class)) %>% 
  pivot_wider(names_from = lc_class, values_from = n, values_fill = 0) %>% 
  mutate(total_area_in_pixels = rowSums(.[,-1]))
```

#7 % canopy tree cover
```{r}
PercentCanopy <- LandCover %>% 
  mutate(PercCanopy = (`Tree Canopy`/total_area_in_pixels)*100) %>% 
  select(CSA2020, PercCanopy)
```

#8 % Impervious
```{r}
PercentLand <- LandCover %>% 
  mutate(Imperv = Structures + `Other Impervious`+ Roads + `Tree Canopy over Structures` + `Tree Canopy over Other Impervious` + `Tree Canopy over Roads` ) %>% 
  mutate(PercentImp = (Imperv/total_area_in_pixels)*100) %>% 
  mutate(Pervious = Water + `Tree Canopy`+ `Scrub\\Shrub`  
         + Herbaceous+ Barren + `Emergent Wetlands`) %>% 
  mutate(PercPervious = (Pervious/total_area_in_pixels)*100) %>% 
  select(CSA2020, PercentImp)
```

#9 Population Density
```{r}
TotalPop <- 
  read_csv('../BaltimoreStreetTreeProject_Large_Data\\Total_Population.csv') %>%
  select(CSA2020, tpop20)

PopDensity <- LandCover %>% 
  left_join(TotalPop) %>%  
  mutate(PopDensity = tpop20/total_area_in_pixels) %>% 
  select(CSA2020, PopDensity)
```

#10 Street Length
```{r}

bmore_bnia <- st_read('../BaltimoreStreetTreeProject_Large_Data\\Community_Statistical_Areas_(CSAs)__Reference_Boundaries/Community_Statistical_Areas_(CSAs)__Reference_Boundaries.shp') |> #BNIA community delineations
  st_make_valid()


roads <- 
  st_read('../BaltimoreStreetTreeProject_Large_Data\\Baltimore Roads/MDOT_Know_Your_Roads.shp'
, as_tibble = TRUE) %>% 
   st_make_valid() |> 
   st_transform(crs = st_crs(bmore_bnia))

NBRoads <- bmore_bnia %>%
  select(CSA2020, Community) %>% 
  st_intersection(roads %>% mutate(road_length_ft = as.double(st_length(.)))) %>% 
  st_drop_geometry() %>% 
  group_by(CSA2020) %>% 
  summarize(sum_road_length_ft = sum(road_length_ft)) %>% 
  mutate(road_length_m = sum_road_length_ft * 0.3048)#convert to meters/SI

```

# join all independent variables 
```{r}
IndVar <- PercentBlk %>% 
  left_join(PercentWhite, by = "CSA2020") %>%
  left_join(Income, by = "CSA2020") %>%
  left_join(Edu, by = "CSA2020") %>%  
  left_join(NBtemperature, by = "CSA2020") %>% 
  left_join(PercVacant, by = "CSA2020") %>% 
  left_join(PercentLand, by = "CSA2020") %>% 
  left_join(PopDensity, by = "CSA2020") %>% 
  left_join(NBRoads, by = "CSA2020") %>% 
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail")
  
write_csv(IndVar,'input_data/Independent_variable.csv')  
```

# Descriptive Statistics
```{r}
summary(IndVar)
```

