---
title: "Tree data"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---

This file contains all dependent variable data for analysis
This file CREATES .CSV WITH COMBINED NB AND STREET TREE ATTRIBUTES

# 0 set up 
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 
library(vegan)
library(codyn)
library(janitor)

getwd()
list.files()
list.files('../BaltimoreStreetTreeProject_Large_Data/')
list.files('../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```

# 1 read in street tree data 
```{r include=FALSE}
street_tree <- 
  st_read('../BaltimoreStreetTreeProject_Large_Data/bc_forestry_trees_20190319/bc_forestry_trees_20190319.shp', as_tibble = TRUE) %>% 
  filter(LOC_TYPE == "Street") 
```

# 2 combine with BNIA neighborhood boundaries
```{r include=FALSE}
bmore_bnia <- st_read('../BaltimoreStreetTreeProject_Large_Data/Community_Statistical_Areas_(2020)/Community_Statistical_Areas__2020_.shp') |> #BNIA community delineations
  st_make_valid() |>
  st_transform(crs = st_crs(street_tree))


  street_tree %>%
  st_join(bmore_bnia) %>%
  st_drop_geometry() |> 
  write_csv(paste0('../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes.csv'))
```

#3 Read in street tree + neighborhood file
```{r}
bnia_street_tree <-
  read_csv('../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes.csv')
```

#4 Clean combined data
```{r}
clean <- 
  bnia_street_tree %>% 
  select(-ID,-OnStr, -OBSERV2, -OBSERV3, -HARD_SCAPE, -PARCELID, -COLLECTOR, -Inv_Time, -INSPECT_DT, -Inspect_TM, -Notes, -UniqueID, -UTILITIES) %>% 
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail") %>% mutate(Species=case_when(
    SPP=='Acer spp.' ~ 'unk999'
    ,SPP=='Amelanchier spp.' ~ 'Amelanchier canadensis'
    ,SPP=='Carya spp.' ~ 'unk999'
    ,SPP=='Chamaecyparis spp.' ~ 'unk999'
    ,SPP=='Cornus spp.' ~ 'unk999'
    ,SPP=='Crataegus spp.' ~ 'Crataegus laevigata'
    ,SPP=='Euonymus spp.'~ 'unknown shrub'
    ,SPP=='Ficus spp.' ~ 'unk999'
    ,SPP=='Fraxinus spp.' ~ 'unk999'
    ,SPP=='Hydrangea spp.' ~ 'unknown shurb'
    ,SPP=='Ilex spp.' ~ 'unk999'
    ,SPP=='Ilex x' ~ 'Ilex x attenuata-Fosteri'
    ,SPP=='Juniperus spp.' ~ 'unk999'
    ,SPP=='Lonicera spp.' ~ 'unknown shrub'
    ,SPP=='Magnolia spp.' ~ 'unk999'
    ,SPP=='Magnolia x' ~ 'Magnolia x soulangiana'
    ,SPP=='Malus spp.' ~ 'Malus sylvestris'
    ,SPP=='Photinia spp.' ~ 'unknown shrub'
    ,SPP=='Picea spp.' ~ 'unk999'
    ,SPP=='Populus spp.' ~ 'unk999'
    ,SPP=="Prunus spp." ~ 'AssignPrunus'
    ,SPP=='Quercus spp.' ~ 'unk999'
    ,SPP=='Quercus x' ~ 'unk999'
    ,SPP=='Rhus spp.' ~ 'unk999'
    ,SPP=='Salix spp.' ~ 'unk999'
    ,SPP=='Taxus spp.' ~ 'Taxus baccata'
    ,SPP=='Thuja spp.' ~ 'Thuja occidentalis'
    ,SPP=='Ulmus spp.' ~ 'unk999'
    ,SPP=='Ulmus x' ~ 'Ulmus hybrid'
    ,SPP=='Viburnum spp.' ~ 'unknown shrub'
    ,SPP=='unknown shrub' ~ 'unknown shrub'
    ,.default = SPP ), .after='SPP') %>% 
  mutate(Species = ifelse(  Species=='AssignPrunus'  # this is to randomly assign prunus
                          , sample(c('Prunus serrulata', 'Prunus subhirtella', 'Prunus x yedoensis')
                                   , size = bnia_street_tree %>% 
                                     filter(SPP=='AssignPrunus') %>% 
                                     nrow()
                                   , replace=T)
                          , Species)) 

```

#5 Filter tree data for only present tree species
```{r include=FALSE}
site_status <- 
  clean %>% 
  mutate(tree_present=
           case_when(
             Species %in% c(  'Vacant Site'
                            , 'Vacant Potential'
                            , 'Stump'
                            , 'Vacant Site Not Suitable'
                            , 'NA'
                            , "Z Add 01"
                            , "Dead") ~ 0
             , CONDITION %in%c('Dead', "Stump", "Absent") ~ 0
             , .default = 1)) #keep dead and stump now?

perc_filled <-
  site_status %>% 
  group_by(CSA2020, tree_present) %>%
  summarise(n= length(Species)) %>% 
  pivot_wider(  id_cols = CSA2020
              , names_from = tree_present
              , values_from = n
              , names_prefix = "T") %>% 
  mutate(  totsites = T1 + T0
         , percent_filled=(T1/totsites)*100)

```

#6 Potential Sites
```{r}
vacant_potential <- 
  clean %>% 
  filter(
      Species=='Vacant Site'
    | Species=='Vacant Potential'
    | Species=='Stump'
    | Species=='NA'
    | Species=="Z Add 01"
    | Species=="Dead") %>% #Stumps and dead no longer considered vacant? 
  mutate(vacant_present = 1) %>%   #for each vacant potential site present column with 1 
  group_by(CSA2020) %>% 
  summarize(sum_vacant_present = sum(vacant_present)) 


```

#7 Total Trees
```{r}
NBTotal <-
  no_vacant %>% 
  group_by(CSA2020) %>% 
  summarize(total_trees =sum(present)) %>% 
  select(total_trees, CSA2020)
```

#8 Percent Filled + Total Sites 
```{r}

total_sites <- 
  vacant_potential %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(all_sites = sum_vacant_present + total_trees) %>% 
  select(CSA2020, all_sites, total_trees)

percfilled <- 
  total_sites %>% 
  mutate(percent_filled = 100*(total_trees/all_sites)) %>% 
  select(CSA2020, percent_filled)

percfilled %>% 
  ggplot(aes(percent_filled) ) +
  geom_density()

total_sites %>% summary()

#histogram
ggplot(total_sites, aes(all_sites)) +
  geom_histogram()

ggplot(total_sites, aes(total_trees)) +
  geom_histogram()

```

#9 Total Abundance by neighborhood
```{r}
#need to exclude unknown trees
NBabund <-
  site_status %>% 
  filter(tree_present == 1) %>% 
  filter(Species != "unk999" & Species!= 'unknown shrub') %>% 
  group_by(CSA2020, Species) %>% 
  summarize(abund=length(Species)) %>% 
  select(CSA2020, Species, abund) 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)
```

#10 Richness and evenness
```{r}
richeven <- community_structure(NBabund, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020")


#norm by total trees
rich_tot <-
  richeven %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(rich_tot = richness/total_trees) %>% 
  select(CSA2020, rich_tot, Evar)
```

#11 Beta Diversity
```{r}
winbeta <- read_csv('../BaltimoreStreetTreeProject/input_data/winbeta.csv')

betadiv <- 
  winbeta %>% 
  select(CSA2020, spave)
```

#12 Average DBH by neighborhood
```{r}
#read in size classes - TF
#created a .csv from .xlsx Nancy sent - TF
#TO DO: PUT CSV INTO GOOGLE DRIVE

missingSPP <- read_csv('../../BaltimoreStreetTreeProject_Large_Data/MissingSpp.csv')

size_classes <- read_csv('../../BaltimoreStreetTreeProject_Large_Data/Size classes.csv') %>% 
  mutate(Species = paste(`Genus Name`, `Species Name`, sep = " " )) %>% 
  select(`Size Class`, Species) %>% 
  rbind(missingSPP)

street_tree_size_classes <- 
  no_vacant %>% 
  left_join(size_classes, by = "Species") 

#list of species with no size class in iTree data
unclassified_species <- 
  street_tree_size_classes %>%
  filter(is.na(`Size Class`)) %>% 
  distinct(Species)

#small trees
AvgDBH_small <- 
  street_tree_size_classes %>% 
  filter(`Size Class` == "Small") %>% 
  select(CSA2020,DBH) %>% 
  mutate(DBH_SI = DBH*2.54) %>%  #convert to SI
  group_by(CSA2020) %>% 
  summarise(AvgDBH_small = mean(DBH_SI), .groups = 'drop')

#medium trees
AvgDBH_med <-
  street_tree_size_classes %>% 
  filter(`Size Class` == "Medium") %>% 
  select(CSA2020,DBH) %>% 
  mutate(DBH_SI = DBH*2.54) %>%  #convert to SI
  group_by(CSA2020) %>% 
  summarise(AvgDBH_med = mean(DBH_SI), .groups = 'drop')

#large trees
AvgDBH_large <- 
  street_tree_size_classes %>% 
  filter(`Size Class` == "Large") %>% 
  select(CSA2020,DBH)%>% 
  mutate(DBH_SI = DBH*2.54) %>%  #convert to SI
  group_by(CSA2020) %>% 
  summarise(AvgDBH_large = mean(DBH_SI), .groups = 'drop')

#OLD - no size classes
# AvgDBH <- no_vacant %>% 
#   select(CSA2020,DBH)%>% 
#   mutate(DBH_SI = DBH*2.54) %>%  #convert to SI
#   group_by(CSA2020) %>% 
#   summarise(AvgDBH = mean(DBH_SI), .groups = 'drop')
  
```

#13 Pit Area Calculation
```{r}

#this is at the tree level, our other calculations are at the nb level
pit_size <- 
  clean %>% 
  mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH)) 
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         ) %>% 
  mutate(  length_SI = length*0.3048 #convert to SI
         , width_SI = width*0.3048 #convert to SI
         , pit_area   = length*width
         ) %>% 
  select(CSA2020, pit_area, X_COORD, Y_COORD, DBH)

ggplot(pit_size, aes(x=pit_area)) + geom_histogram()

cor.test(pit_size$pit_area, pit_size$DBH)
# r=0.342

large_pit <- 
  pit_size %>% 
  mutate(size_threshhold = ifelse(pit_area >= 2.97, "greater", "less" )) %>% 
  group_by(CSA2020, size_threshhold) %>% 
  summarize(n = length(size_threshhold)) %>% 
  pivot_wider(names_from = size_threshhold, values_from = n, values_fill = 0) %>% 
  left_join(total_sites, by = "CSA2020") %>% 
  mutate(Perc2.97 = greater / all_sites*100) %>% 
  select(CSA2020, Perc2.97)


```

#14 Tree Condition
```{r}
#sum for each condition in a neighborhood
#pivot
condition <- 
  clean %>%
  group_by(CSA2020, CONDITION) %>% 
  summarize(n=length(CONDITION)) %>% 
  pivot_wider(names_from = CONDITION, values_from = n, values_fill = 0) %>% 
  select(CSA2020, Dead ,Fair ,Good, Poor ,Stump)

# numeric condition
condition_grade <- 
  no_vacant %>% 
  filter(CONDITION!="Sprout" & CONDITION!="Sprouts") %>% 
  mutate(grade = 
           case_when(
               CONDITION == "Poor" ~ 1
             , CONDITION == "Fair" ~ 2
             , CONDITION == "Good" ~ 3
    )) %>% 
  select(CSA2020,grade) 


```

#15 Condition and Pit Area
```{r}

no_vacant |> 
  filter(CONDITION != "Sprout" & CONDITION != "Sprouts") |>
   mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
          , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
          , pit_area   = length*width
          ) |>
  ggplot(aes(CONDITION, pit_area)) +
  geom_boxplot()
  
```

#16 % good/poor trees
```{r}
perc_condition <- 
  NBTotal %>% 
  left_join(condition, by = "CSA2020") %>% 
  mutate(PercGood = 100*(Good/total_trees)) %>% 
  mutate(PercPoor = 100*((Poor)/total_trees)) %>% 
  mutate(PercDead = 100*((Dead + Stump/total_trees))) %>% 
  select(CSA2020, PercPoor, PercGood, PercDead)

```

#17 Combine all dependent variables
```{r}
 DepVar <- percfilled %>% 
  left_join(total_sites) %>%  
  left_join(rich_tot, by = "CSA2020") %>% 
  left_join(betadiv, by = "CSA2020") %>% 
  
  left_join(AvgDBH_small, by = "CSA2020") %>% 
  left_join(AvgDBH_med, by = "CSA2020") %>% 
  left_join(AvgDBH_large, by = "CSA2020") %>% 
  
  left_join(perc_condition, by = "CSA2020") %>% 
  left_join(large_pit, by = "CSA2020")

  
write_csv(DepVar,'../input_data/Dependent_variable.csv')  
```

#Descriptive Statistics
```{r}
summary(DepVar)
```

