---
title: "Analysis Final"
output:
  pdf_document: default
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---

#0 set up 
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 

library(grid)
library(gridExtra)

library(cowplot)
#working directory
getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)

var_names <- as_labeller(
     c(`PercBlk` = "% Black", `PercWhite` = "% White",`mhhi20` = "Income", 
       `bahigher20` = "Bachelor's or higher", `avg_temp` = "Average Temperature"))

var_names_2 <- as_labeller(
     c(`AvgDBH` = "Average DBH", `Evar` = "Evenness",`Perc32` = "Ideal Pits", 
       `PercNat` = "Nativity", `PercGood` = "% Good", `PercPoor` = "% Poor", `percent_filled` = "% Filled", `richness` = "Richness"))
```

#1 Combined dep+ind
```{r include=FALSE}
Dependent <- read_csv('../input_data/Dependent_variable.csv')
Independent<- read_csv('../input_data/Independent_variable.csv')

# Dexter path
Dependent <- read_csv('input_data/Dependent_variable.csv')
Independent<- read_csv('input_data/Independent_variable.csv')

Combined <- Dependent %>% 
  left_join(Independent, by = "CSA2020") 

```

#2 All p values and correlations
```{r include=FALSE}
 
#empty df to store results
cor_results <- data.frame()

# Get names of all variables in df

Vars <- Combined %>% 
  select(where(is.numeric)) 

variables <- names(Vars)

# Loop through all combinations of variables
for (i in 1:(length(variables) - 1)) {
  for (j in (i + 1):length(variables)) {
    
    # Extract the two variables for the current combination
    var1 <- Vars[[variables[i]]]
    var2 <- Vars[[variables[j]]]
    
    # Run correlation test
    cor_test_result <- cor.test(var1, var2)
    
    # Extract correlation coefficient and p-value
    correlation_coefficient <- cor_test_result$estimate
    p_value <- cor_test_result$p.value
    
    # Create a row for the results dataframe
    result_row <- data.frame(
      Variable1 = variables[i],
      Variable2 = variables[j],
      CorrelationCoefficient = correlation_coefficient,
      PValue = p_value
    )
    
    # Bind row to results df
    cor_results <- rbind(cor_results, result_row)
  }
}

# View the resulting dataframe
print(cor_results)

# Different Method
# # like this
# correlation::correlation(Vars)
# 
# # # or like this,
# pvalues <- 
# Combined |> 
#   select(where(is.numeric)) |> 
#   correlation::correlation() |>
#   tibble() # can add this 
```

#3 Filtered p < 0.05
```{r message=FALSE, include=FALSE}

p_filtered_0.05 <- cor_results %>% 
  filter(PValue<0.05) 

```

# Filtered p < 0.001
```{r message=FALSE, warning=FALSE, include=FALSE}
p_filtered_0.001 <- cor_results %>% 
  filter(PValue<0.001)

```

#4 Total Tree Pits
```{r}
totpits_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "all_sites" | Variable2 == "all_sites")

view(totpits_r)

totpits_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "all_sites" | Variable2 == "all_sites")

view(totpits_r2)
```

#5 Stocking
```{r echo=FALSE, message=FALSE, warning=FALSE}
stocking_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "percent_filled" | Variable2 == "percent_filled")

view(stocking_r)

#perc_filled_plot <-
  Combined |> 
  tidylog::select(
                   percent_filled  # depend
                  , bahigher20
                  ) |> 
  pivot_longer(-percent_filled) |> 
  ggplot(aes(value, percent_filled)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 1, nrow =1
             , labeller = var_names
             ) +
  labs(
         x = 'Predictor'
       , y = 'Percent Filled'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

# <0.001

stocking_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "percent_filled" | Variable2 == "percent_filled")

view(stocking_r2)  
```

#6 Richness
```{r}
richtot_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "rich_tot" | Variable2 == "rich_tot")

view(richtot_r)

richtot_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "rich_tot" | Variable2 == "rich_tot")

view(richtot_r2)
```

#7 Evenness
```{r echo=FALSE, message=FALSE, warning=FALSE}
even_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Evar" | Variable2 == "Evar")

view(even_r)

#p<0.001  
even_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "Evar" | Variable2 == "Evar")

view(even_r2)

```

#8 Beta Diversity
```{r}
betadiv_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "spave" | Variable2 == "spave")

view(betadiv_r)

betadiv_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "spave" | Variable2 == "spave")

view(betadiv_r2)
```

#9 Average DBH
```{r echo=FALSE, message=FALSE, warning=FALSE}
dbh_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "AvgDBH" | Variable2 == "AvgDBH")

view(dbh_r)

dbh_plot <- Combined |> 
  tidylog::select(
                    AvgDBH  # depend 
                  , PercBlk
                  , avg_temp
                  ) |> 
  pivot_longer(-AvgDBH) |> 
  ggplot(aes(value, AvgDBH)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Average DBH'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

dbh_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "AvgDBH" | Variable2 == "AvgDBH")

view(dbh_r2)
```

#10 Poor Condition
```{r echo=FALSE, message=FALSE, warning=FALSE}
poor_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercPoor" | Variable2 == "PercPoor")

view(poor_r)

poor_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "PercPoor" | Variable2 == "PercPoor")

view(poor_r2)
```

#11 Good Condition
```{r echo=FALSE, message=FALSE, warning=FALSE}
good_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercGood" | Variable2 == "PercGood")

view(good_r)

good_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "PercGood" | Variable2 == "PercGood")

view(good_r2)
```

#12 Target Pit Size
```{r echo=FALSE, message=FALSE, warning=FALSE}
perc32_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Perc32" | Variable2 == "Perc32")

view(perc32_r)

# p<0.001
perc32_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "Perc32" | Variable2 == "Perc32")

view(perc32_r2)

```

#13 % Flowering 
```{r}
percflor_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercFlor" | Variable2 == "PercFlor")

view(percflor_r)

percflor_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "PercFlor" | Variable2 == "PercFlor")

view(percflor_r2)
```

#14 % Showy
```{r}
percshow_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercShowy" | Variable2 == "PercShowy")

view(percshow_r)

percshow_r2 <- p_filtered_0.001 %>% 
  filter(Variable1 == "PercShowy" | Variable2 == "PercShowy")

view(percshow_r2)
```

#15 Average Showiness
```{r}
avgshow_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "AvgShowy" | Variable2 == "AvgShowy")

view(avgshow_r)


```

#18 Avg showy
```{r}
avgshowy_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "avgShowy" | Variable2 == "AvgShowy")

view(avgshowy_r)
```


# r value matrix
```{r}
cor_matrix <- cor_results %>% 
  select(Variable1, Variable2, CorrelationCoefficient) %>% 
  spread(Variable2, CorrelationCoefficient) %>% 
  column_to_rownames(var = "Variable1")

print(cor_matrix)


write_xlsx(cor_matrix, '..\\..\\BaltimoreStreetTreeProject_Large_Data\\Rmatrix.xlsx')

```

# scatter plots
```{r}
# stocking vs % black
A<- ggplot(Combined, aes(x = bahigher20, y = percent_filled))+
  geom_point(color = "#00e293", size=2.5)+
  labs(x = "Bachelor's", y = "Stocking") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))

#%poor vs % black
B <- ggplot(Combined, aes(x = PercBlk, y = PercPoor))+
  geom_point(color = "#a55aa4", size=2.5)+
  labs(x = "% Black", y = "Poor Condition") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))

#richness vs temperature
C <- ggplot(Combined, aes(x = avg_temp, y = rich_tot))+
  geom_point(color = "#ff794c", size=2.5)+
  labs(x = "Temperature (C)", y = "Relative Richness") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))
#richness vs % black
D <- ggplot(Combined, aes(x = PercBlk, y = rich_tot))+
  geom_point(color = "#ff794c", size=2.5)+
  labs(x = "% Black", y = "Richness") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))
#DBH vs temp
E <- ggplot(Combined, aes(x = avg_temp, y = AvgDBH))+
  geom_point(color = "#00afff", size=2.5)+
  labs(x = "Temperature (C)", y = "Average DBH") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))

#showy vs %white
G <- ggplot(Combined, aes(x = PercWhite, y = PercShowy))+
  geom_point(color = "#e95789", size=2.5)+
  labs(x = "% White", y = "% Showy") + 
  theme_bw()+
    theme(panel.grid=element_blank(),
        text=element_text(size=15))
```

# arrange in grid
```{r}
library(gtable)

Fig_grid <- plot_grid(A, B, C, D,E, G, labels = c('A', 'B', 'C', 'D', 'E', 'F'), ncol = 2,vjust =1)


ggsave("scatterplot grid.jpg",width=180, height=200, unit="mm", plot=Fig_grid, dpi=300 )
```

