---
title: "Analysis"
author: "Street Tree"
date: "`r format(Sys.Date())`"
output:
  pdf_document: default
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---
#0 set up
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 

#working directory
getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```

This file will conduct correlational analyses on combined data of demographics and street tree variables 

#1 Combine dep+ind
```{r include=FALSE}
# Dependent <- read_csv('../../BaltimoreStreetTreeProject/Dependent_variable.csv')
# Independent<- read_csv('../../BaltimoreStreetTreeProject/Independent_variable.csv') 

Dependent <- read_csv('../input_data/Dependent_variable.csv')
Independent<- read_csv('../input_data/Independent_variable.csv')

Combined <- Dependent %>% 
  left_join(Independent, by = "CSA2020") 

  

Combined |> glimpse()
```


#2 All correlations
```{r include=FALSE}

(cors_tibble <- 
  Combined |> 
  select(where(is.numeric)) |> # just numeric variables
  cor(use = 'pairwise.complete') |> # instead of drop_na
  data.frame() |> # changing from matrix to dataframe
  rownames_to_column(var = 'var1') |> 
  as_tibble() |> 
  pivot_longer(-var1, names_to = 'var2') |> # make long
  filter(abs(value) !=1) # drop perfect correlations
 ) 

#cors_tibble |> View() # for interactivity

# per group, most correlated
 cors_tibble |> 
   group_by(var1) |> 
   arrange(desc(value)) |> 
   slice(1:5) |> # top five most correlated
   ungroup() #|> View()

```

#3 Sort correlations for r>= |0.4|
```{r include=FALSE}

# filter for correlations higher than .9 (values higher than .89 not correlated with demographics)
cor_0.9 <- cors_tibble %>% 
  filter(value >= 0.9) %>%  
  group_by(var1) 

#filter for correlations r = 0.8-0.89
cor_0.8 <- cors_tibble %>% 
  filter(value >=0.8&value < .9) %>%  
  group_by(var1)

#filter for var1= predicted independent variables and var2 = dependent variables, |r|>= 0.3
cor_filtered <- cors_tibble %>% 
  #filter var1 = independent variables
  filter(var1 =='PercBlk'|var1 =='PercWhite'|
           var1=='mhhi20'|var1=='bahigher19'|
           var1=='avg_temp')%>%
  #filter var2 = dependent variables
  filter(var2=='richness'|var2=='Evar'|var2=='AvgDBH'|var2=='cv'|
           var2=='richness_SM'|var2=='Evar_SM'|var2=='sm_rich_tot'|
           var2=='richness_L'|var2=='Evar_L'|var2=='L_rich_tot'|
           var2=='PercNat'|var2=='avg_temp'|var2=='percent_filled'|var2== 'all_sites'|
           var2== 'PercGood'|var2== 'PercPoor'|var2 == 'PercSmall'|var2 == 'PercMedium'|var2 == 'PercLarge') %>%
  filter(value>=0.4 | value<=-0.4)  
```

#4 First pass scatter plots
```{r echo=FALSE}

Combined |> glimpse()



#richness norm by total trees
Combined |> 
  tidylog::select(#CSA2020 # ID
                   richness_SM # depend
                  , mhhi20 # independents, income
                  , PercWhite # race
                  , PercBlk
                  , bahigher19 # edu
                  , avg_temp
                  ) |> 
  pivot_longer(-richness_SM) |> 
  ggplot(aes(value,richness_SM)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  geom_smooth(col = 'pink') + 
  facet_wrap(~name, scales = 'free_x'
             , ncol = 3
             ) +
  labs(title = 'Total Number of Trees'
       , subtitle = 'Normalized by Neighborhood Area'
       , captions = 'sources'
       , x = 'predictor'
       , y = 'my outcome of interest (total street trees in this case)'
       ) + 
  theme_bw(14) +
  NULL


```

Scatter plots for correlations that met criteria r>=|0.4| for indep vs dep variables:

# 5 Evenness  
```{r echo=FALSE, message=FALSE, warning=FALSE}
Combined |> 
  tidylog::select(
                   Evar  # depend
                  , avg_temp
                  ) |> 
  pivot_longer(-Evar) |> 
  ggplot(aes(value, Evar)) +
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  labs(x = 'Average Temperature'
       , y = 'Evenness' 
       ) + theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$avg_temp, Combined$Evar)

#dataset with no outliers

evenness <- Combined %>% 
  filter(Evar < 0.6) %>% 
  select(Evar  # depend
                  , avg_temp)

  ggplot(evenness, aes(avg_temp, Evar)) +
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  labs(x = 'Average Temperature'
       , y = 'Evenness' 
       ) + theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

cor.test(evenness$avg_temp, evenness$Evar)
```

# 6 Richness SM
```{r echo=FALSE, message=FALSE, warning=FALSE}

var_names <- as_labeller(
     c(`PercBlk` = "% Black", `PercWhite` = "% White",`mhhi20` = "Median Household Income", 
       `bahigher19` = "Bachelor's or higher"))

Combined |> 
  tidylog::select(
                    richness_SM  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher19
                  ) |> 
  pivot_longer(-richness_SM) |> 
  ggplot(aes(value, richness_SM)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'SM Trees Richness'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

cor.test(Combined$PercBlk, Combined$richness_SM)
cor.test(Combined$PercWhite, Combined$richness_SM)
cor.test(Combined$mhhi20, Combined$richness_SM)
cor.test(Combined$bahigher19, Combined$richness_SM)
```

# 7 Norm S Richness  
```{r echo=FALSE, message=FALSE, warning=FALSE}

#normalized by total number of trees
Combined |> 
  tidylog::select(
                    sm_rich_tot  # depend
                  , avg_temp
                  ) |>
  pivot_longer(-sm_rich_tot) |> 
  ggplot(aes(value, sm_rich_tot)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(
        x = 'Average Temperature'
       , y = 'SM Trees Richness'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))+
  NULL

#dataset with no outliers

norm_rich_SM <- Combined %>% 
  filter(sm_rich_tot< 0.1) %>% 
  select(sm_rich_tot  # depend
                  , avg_temp)

ggplot(norm_rich_SM, aes(avg_temp, sm_rich_tot)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(
        x = 'Average Temperature'
       , y = 'SM Trees Richness'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))+
  NULL

cor.test(norm_rich_SM$avg_temp, norm_rich_SM$sm_rich_tot)
```

# 8 SM Evenness
```{r echo=FALSE, message=FALSE, warning=FALSE}
Combined |> 
  tidylog::select(
                   Evar_SM  # depend
                  , avg_temp
                  ) |> 
  pivot_longer(-Evar_SM) |> 
  ggplot(aes(value, Evar_SM)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(title = 'Small Trees Evenness'
       , x = 'Average Temperature'
       , y = 'Richness'
       ) + 
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5))
  NULL

cor.test(Combined$avg_temp, Combined$Evar_SM)
```

# 9 Richness L
```{r echo=FALSE, message=FALSE, warning=FALSE}
Combined |> 
  tidylog::select(
                   richness_L  # depend
                  , avg_temp
                  ) |> 
  pivot_longer(-richness_L) |> 
  ggplot(aes(value, richness_L)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(title = 'Large Trees Richness'
       , x = 'Average Temperature'
       , y = 'L Richness'
       ) + 
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5))
  NULL

cor.test(Combined$avg_temp, Combined$richness_L)
```

# 10 Richness L Normalized
```{r echo=FALSE, message=FALSE, warning=FALSE}

#normalized by total number of trees
Combined |> 
  tidylog::select(
                   L_rich_tot  # depend
                  , avg_temp
                  
                  ) |> 
  pivot_longer(-L_rich_tot) |> 
  ggplot(aes(value, L_rich_tot)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(title = 'Large Trees Richness'
       , subtitle = 'Normalized by Total Trees'
       , x = 'Predictor'
       , y = 'L Richness'
       ) + 
  theme_bw(12) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.subtitle = element_text(hjust = 0.5))
  NULL

cor.test(Combined$avg_temp, Combined$L_rich_tot)

```

# 11 Avg DBH
```{r echo=FALSE, message=FALSE, warning=FALSE}

Combined |> 
  tidylog::select(
                   AvgDBH  # depend
                  , avg_temp
                  ) |> 
  pivot_longer(-AvgDBH) |> 
  ggplot(aes(value, AvgDBH)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs( x = 'Average Temperature'
       , y = 'Average DBH'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$avg_temp, Combined$AvgDBH)

```

# 12 Total Sites (Filled)
```{r echo=FALSE}

Perc_names <- as_labeller(
     c(`PercBlk` = "Percent Black", `PercWhite` = "Percent White", 
       `bahigher19` = "Bachelor's or higher", `total_trees` = "Total Trees"))
Combined |> 
  tidylog::select(
                   percent_filled  # depend
                  , PercBlk
                  , PercWhite
                  , bahigher19
                  , total_trees
                  ) |> 
  pivot_longer(-percent_filled) |> 
  ggplot(aes(value, percent_filled)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, nrow =2
             , labeller = Perc_names
             ) +
  labs(
         x = 'Predictor'
       , y = 'Percent Filled'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$percent_filled, Combined$PercBlk)
cor.test(Combined$percent_filled, Combined$PercWhite)
cor.test(Combined$percent_filled, Combined$bahigher19)
cor.test(Combined$percent_filled, Combined$mhhi20)
```

# 13 % Filled Sites with Small Trees
```{r}

```

# 14 % filled Sites with Large Trees
```{r}

```

# 15 Percent Nativity
```{r echo=FALSE, message=FALSE, warning=FALSE}
Combined |> 
  tidylog::select(
                   PercNat  # depend
                  , avg_temp
                  ) |> 
  pivot_longer(-PercNat) |> 
  ggplot(aes(PercNat, value)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  # facet_wrap(~name, scales = 'free_x'
  #            , ncol = 3
  #            ) +
  labs(x = 'Average Temperature'
       , y = 'Percent Native'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$avg_temp, Combined$PercNat)


```

# 16 Total Sites 
```{r}
var_names <- as_labeller(
     c(`PercBlk` = "% Black", `PercWhite` = "% White",`mhhi20` = "Median Household Income", 
       `bahigher19` = "Bachelor's or higher"))

Combined |> 
  tidylog::select(
                    all_sites  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher19
                  ) |> 
  pivot_longer(-all_sites) |> 
  ggplot(aes(value, all_sites)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Potential Tree Sites'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$PercBlk, Combined$all_sites)
cor.test(Combined$PercWhite, Combined$all_sites)
cor.test(Combined$mhhi20, Combined$all_sites)
cor.test(Combined$bahigher19, Combined$all_sites) 
```

# 17 Percent Condition
```{r}
var_names <- as_labeller(
     c(`PercBlk` = "% Black", `PercWhite` = "% White",`mhhi20` = "Median Household Income", 
       `bahigher19` = "Bachelor's or higher"))

Combined |> 
  tidylog::select(
                    PercGood  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher19
                  ) |> 
  pivot_longer(-PercGood) |> 
  ggplot(aes(value, PercGood)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Trees in Good Condition'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL
  
cor.test(Combined$PercBlk, Combined$PercGood)
cor.test(Combined$PercWhite, Combined$PercGood)
cor.test(Combined$mhhi20, Combined$PercGood)
cor.test(Combined$bahigher19, Combined$PercGood)


Combined |> 
  tidylog::select(
                    PercPoor  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher19
                  ) |> 
  pivot_longer(-PercPoor) |> 
  ggplot(aes(value, PercPoor)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Trees in Good Condition'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL

cor.test(Combined$PercBlk, Combined$PercPoor)
cor.test(Combined$PercWhite, Combined$PercPoor)
cor.test(Combined$mhhi20, Combined$PercPoor)
cor.test(Combined$bahigher19, Combined$PercPoor)
```

# 18 SM Richness no outlier
```{r}
# no outlier
filled_richness <- Combined %>% 
  filter(sm_rich_tot < 0.1) %>% 
  select(sm_rich_tot,CSA2020,percent_filled)


ggplot(filled_richness, aes(percent_filled, sm_rich_tot)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  labs(
        x = 'Percent Filled'
       , y = 'SM Trees Richness'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
 NULL

(cor.test(filled_richness$percent_filled, filled_richness$sm_rich_tot))

 
```

# 19 Percent Canopy (as dependent)
```{r}
  Combined |> 
  tidylog::select(
                    PercCanopy  # depend 
                  , percent_filled 
                  , all_sites
                  , avg_temp #indep
                  ) |> 
  pivot_longer(-PercCanopy) |> 
  ggplot(aes(value, PercCanopy)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Canopy'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20))
  NULL
  
(cor.test(Combined$percent_filled, Combined$PercCanopy))
(cor.test(Combined$all_sites, Combined$PercCanopy))
(cor.test(Combined$avg_temp, Combined$PercCanopy))
```

