---
title: "Tree data"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---

This file contains all dependent variable data for analysis

# 0 set up 
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 
library(dplyr)
library(vegan)
library(codyn)

getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```

# 1 read in street tree data 
```{r include=FALSE}
street_tree <- 
  st_read('../../BaltimoreStreetTreeProject_Large_Data\\bc_forestry_trees_20190319\\bc_forestry_trees_20190319.shp'
, as_tibble = TRUE) %>% 
  filter(LOC_TYPE == "Street") 
```

# 2 combine with BNIA neighborhood boundaries
```{r include=FALSE}
bmore_bnia <- st_read('../../BaltimoreStreetTreeProject_Large_Data\\Community_Statistical_Areas(2020)/Community_Statistical_Areas__2020_.shp') |> #BNIA community delineations
  st_make_valid() |>
  st_transform(crs = st_crs(street_tree))


  street_tree %>%
  st_join(bmore_bnia) %>%
  st_drop_geometry() |> 
  write_csv(paste0('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_', Sys.Date(), '.csv'))
```

#3 Read in street tree + neighborhood file
```{r include=FALSE}
bnia_street_tree <-
  read_csv('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_2023-10-31.csv')
```

#4 Clean combined data
```{r}
clean <- bnia_street_tree %>% 
  select(-ID,-OnStr, -OBSERV2, -OBSERV3, -HARD_SCAPE, -PARCELID, -COLLECTOR, -Inv_Time, -INSPECT_DT, -Inspect_TM, -Notes, -UniqueID, -UTILITIES) %>% 
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail")
```

#5 Filter tree data for only present tree species
```{r include=FALSE}
no_vacant <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species!='Vacant Site'&Species!='Vacant Potential'&Species!='Stump'&Species!='Vacant Site Not Suitable'&Species!='NA'&Species!='unknown shrub'&Species!="Z Add 01"&Species!=" "&Species!="Dead")%>%
  filter(CONDITION!="Dead"&CONDITION!="Stump"&CONDITION!="Absent")%>% 
  mutate(present=1) #for each tree make a present column with 1

```

#6 Unfilled Sites
```{r}
vacant_potential <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species=='Vacant Site'|Species=='Vacant Potential'|Species=='Stump'|Species=='NA'|Species=='unknown shrub'|Species=="Z Add 01"|Species=="Dead")%>% 
  mutate(vacant_present=1) %>%   #for each vacant potential site present column with 1 
  group_by(CSA2020) %>% 
  summarize(sum_vacant_present =sum(vacant_present)) 


```

#7 Total Trees
```{r}
NBTotal <-no_vacant %>% 
  group_by(CSA2020) %>% 
  summarize(total_trees =sum(present)) %>% 
  select(total_trees, CSA2020)
```

#8 Percent Filled + Total Sites 
```{r}

total_sites <- vacant_potential %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(all_sites = sum_vacant_present + total_trees) %>% 
  select(CSA2020, all_sites, total_trees)

percfilled <- total_sites %>% 
  mutate(percent_filled = 100*(total_trees/all_sites)) %>% 
  select(CSA2020, percent_filled)

percfilled %>% 
  ggplot(aes(percent_filled) ) +
  geom_density()

total_sites %>% summary()

#histogram
ggplot(total_sites, aes(all_sites)) +
  geom_histogram()

ggplot(total_sites, aes(total_trees)) +
  geom_histogram()

```

#9 Total Abundance by neighborhood
```{r}
#need to exclude unknown trees
NBabund <-no_vacant %>% 
  filter(Species != "unknown tree") %>% 
  group_by(CSA2020, Species) %>% 
  summarize(abund=sum(present)) %>% 
  select(CSA2020, Species, abund) 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)


```

#10 Richness and evenness
```{r}
richeven <- community_structure(NBabund, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020")


#norm by total trees
rich_tot <- richeven %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(rich_tot = richness/total_trees) %>% 
  select(CSA2020, rich_tot)
```

#11 Beta Diversity
```{r}
winbeta <- read_csv('../../BaltimoreStreetTreeProject/input_data/winbeta.csv')

betadiv <- winbeta %>% 
  select(CSA2020, spave)
```

#12 Average DBH by neighborhood
```{r}
AvgDBH <- no_vacant %>% 
  select(CSA2020,DBH)%>% 
  group_by(CSA2020) %>% 
  summarise(AvgDBH = mean(DBH), .groups = 'drop')

#convert conditions to numeric 
```

#14 Pit Area Calculation
```{r}

#this is at the tree level, our other calculations are at the nb level
pit_size <- clean %>% 
  mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         , pit_area   = length*width
         ) %>% 
  select(CSA2020, pit_area, X_COORD, Y_COORD, DBH)

ggplot(pit_size, aes(x=pit_area)) + geom_histogram()

cor.test(pit_size$pit_area, pit_size$DBH)
# r=0.342

large_pit <- pit_size %>% 
  mutate(size_threshhold = ifelse(pit_area >= 32, "great32", "less32" )) %>% 
  group_by(CSA2020, size_threshhold) %>% 
  summarize(n=length(size_threshhold)) %>% 
  pivot_wider(names_from = size_threshhold, values_from = n, values_fill = 0) %>% 
  left_join(total_sites, by = "CSA2020") %>% 
  mutate(Perc32 = great32/all_sites*100) %>% 
  select(CSA2020, Perc32)


```

#15 Tree Condition
```{r}
#sum for each condition in a neighborhood
#pivot
condition <- no_vacant %>%
  group_by(CSA2020, CONDITION) %>% 
  summarize(n=length(CONDITION)) %>% 
  pivot_wider(names_from = CONDITION, values_from = n, values_fill = 0)

# numeric condition
condition_grade <- no_vacant %>% 
  filter(CONDITION!="Sprout"&CONDITION!="Sprouts") %>% 
  mutate(grade = 
  case_when(
    CONDITION == "Poor" ~ 1,
    CONDITION == "Fair" ~ 2,
    CONDITION == "Good" ~ 3,
    )) %>% 
  select(CSA2020,grade) 


#Condition vs pit size
#%>% 
  # left_join(no_vacant_pit, by = "CSA2020")
```

#16 Condition and Pit Area
```{r}

no_vacant |> 
  filter(CONDITION!="Sprout"&CONDITION!="Sprouts") |>
   mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         , pit_area   = length*width
         ) |>
  ggplot(aes(CONDITION, pit_area)) +
  geom_boxplot()
  
```

#17 % good/poor trees
```{r}
perc_condition <- NBTotal %>% 
  left_join(condition, by = "CSA2020") %>% 
  mutate(PercGood = 100*(Good/total_trees)) %>% 
  mutate(PercPoor = 100*(Poor/total_trees)) %>% 
  select(CSA2020, PercPoor, PercGood)
```

#18 Flowering Trees
```{r}
# read in flowering data
flowering <- read_csv("../../BaltimoreStreetTreeProject_Large_Data/Flowering Data.csv", show_col_types = FALSE)

#select for flowering yes/no
flower_y_n <- flowering %>% 
  select(CSA2020, Flowering) %>% 
  filter(Flowering =="yes" | Flowering == "no")


PercFlower <- flower_y_n %>%
  group_by(CSA2020, Flowering) %>% 
  summarize(n=length(Flowering)) %>%
  pivot_wider(names_from = Flowering, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(PercFlor = (yes/total_trees)*100) %>% 
  select(CSA2020, PercFlor)


```

#19 Showy
```{r}
#select for showy = 3 (most showy)
showiness <- flowering %>% 
  select(CSA2020, Visibility)

PercShowy <- showiness %>% 
  group_by(CSA2020, Visibility) %>% 
  summarize(n=length(Visibility)) %>%
  pivot_wider(names_from = Visibility, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  rename("showy" = '2') %>% 
  mutate(PercShowy = (showy/total_trees)*100) %>% 
  select(CSA2020, PercShowy)
```

#20 Color abundance
```{r}

color_abund <- flowering %>% 
  filter(Visibility>=2) %>% 
  group_by(CSA2020, Color) %>% 
  summarize(abund=length(Color))  


```

#21 Color Richness
```{r}
color_rich <- community_structure(color_abund, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020") %>% 
  rename(color_rich = richness) %>% 
  rename(color_evar = Evar)
```


#22 Average Showiness
```{r}

avgShowy <- showiness %>% 
  filter(Visibility != "NA") %>% 
  group_by(CSA2020) %>% 
  summarise(AvgShowy = mean(Visibility), .groups = 'drop')

```

#27 5 Hottest NB Top 5 Species 
```{r}
small_trees_abund <- no_vacant %>%
  filter(DBH>=0&DBH<=5) %>%  
  group_by(CSA2020, Species) %>% 
  summarize(n=length(Species)) %>%
  pivot_wider(names_from = Species, values_from = n, values_fill = 0)

small_abund_temp <- NBtemperature %>%
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail") %>% 
  left_join(small_trees_abund, by = "CSA2020") %>% 
  arrange(desc(avg_temp)) %>% 
  select(-avg_temp)

#top species for Madison
madison_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Madison/East End") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(madison_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Madison/East End"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Patterson Park
patterson_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Patterson Park North & East") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(patterson_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Patterson Park North & East"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# oldtown
oldtown_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Oldtown/Middle East") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(oldtown_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Oldtown/Middle East"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Highlandtown
highland_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Highlandtown") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(highland_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Highlandtown"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Downtown/Seton Hill
downtown_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Downtown/Seton Hill") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(downtown_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Downtown/Seton Hill"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#28 5 Coolest NB Top 5 Species
```{r}
#mount washington
mtwash_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Mount Washington/Coldspring") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(mtwash_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Mount Washington/Coldspring"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#forest park
forpark_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Forest Park/Walbrook") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(forpark_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Forest Park/Walbrook"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#edmonson village
edmonson_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Edmonson Village") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(edmonson_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Edmonson Village"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#greater roland park
roland_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Greater Roland Park/Poplar Hill") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(roland_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Greater Roland Park/Poplar Hill"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#beechfield
beechfield_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Beechfield/Ten Hills/West Hills") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(beechfield_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Beechfield/Ten Hills/West Hills"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#29 Combine all dependent variables
```{r}
 DepVar <- percfilled %>% 
  left_join(total_sites) %>% 
  left_join(richeven, by = "CSA2020") %>% 
  left_join(rich_tot, by = "CSA2020") %>% 
  left_join(betadiv, by = "CSA2020") %>% 
  
  left_join(AvgDBH, by = "CSA2020") %>% 
  
  left_join(perc_condition, by = "CSA2020") %>% 
  left_join(large_pit, by = "CSA2020") %>% 
  
  left_join(PercFlower, by = "CSA2020") %>% 
  left_join(PercShowy, by = "CSA2020") %>%  
  left_join(color_rich, by = "CSA2020") %>% 
  left_join(avgShowy, by = "CSA2020")

  
write_csv(DepVar,'../input_data/Dependent_variable.csv')  
```




