---
title: "Tree data"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---

This file contains all dependent variable data for analysis

# 0 set up 
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 
library(dplyr)
library(vegan)
library(codyn)

getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```

# 1 read in street tree data 
```{r include=FALSE}
street_tree <- 
  st_read('../../BaltimoreStreetTreeProject_Large_Data\\bc_forestry_trees_20190319\\bc_forestry_trees_20190319.shp'
, as_tibble = TRUE) %>% 
  filter(LOC_TYPE == "Street") 
```

# 2 combine with BNIA neighborhood boundaries
```{r include=FALSE}
bmore_bnia <- st_read('../../BaltimoreStreetTreeProject_Large_Data\\Community_Statistical_Areas(2020)/Community_Statistical_Areas__2020_.shp') |> #BNIA community delineations
  st_make_valid() |>
  st_transform(crs = st_crs(street_tree))


  street_tree %>%
  st_join(bmore_bnia) %>%
  st_drop_geometry() |> 
  write_csv(paste0('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_', Sys.Date(), '.csv'))
```

#3 Read in street tree + neighborhood file
```{r include=FALSE}
bnia_street_tree <-
  read_csv('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_2023-10-31.csv')
```

#4 Clean combined data
```{r}
clean <- bnia_street_tree %>% 
  select(-ID,-OnStr, -OBSERV2, -OBSERV3, -HARD_SCAPE, -PARCELID, -COLLECTOR, -Inv_Time, -INSPECT_DT, -Inspect_TM, -Notes, -UniqueID, -UTILITIES) %>% 
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail")
```

#5 Filter tree data for only present tree species
```{r include=FALSE}
no_vacant <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species!='Vacant Site'&Species!='Vacant Potential'&Species!='Stump'&Species!='Vacant Site Not Suitable'&Species!='NA'&Species!='unknown shrub'&Species!="Z Add 01"&Species!=" "&Species!="Dead")%>%
  filter(CONDITION!="Dead"&CONDITION!="Stump"&CONDITION!="Absent")%>% 
  mutate(present=1) #for each tree make a present column with 1

```

#6 Unfilled Sites
```{r}
vacant_potential <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species=='Vacant Site'|Species=='Vacant Potential'|Species=='Stump'|Species=='NA'|Species=='unknown shrub'|Species=="Z Add 01"|Species=="Dead")%>% 
  mutate(vacant_present=1) %>%   #for each vacant potential site present column with 1 
  group_by(CSA2020) %>% 
  summarize(sum_vacant_present =sum(vacant_present)) 


```

#7 Total Trees
```{r}
NBTotal <-no_vacant %>% 
  group_by(CSA2020) %>% 
  summarize(total_trees =sum(present)) %>% 
  select(total_trees, CSA2020)
```

#8 Percent Filled + Total Sites 
```{r}

total_sites <- vacant_potential %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(all_sites = sum_vacant_present + total_trees) %>% 
  select(CSA2020, all_sites, total_trees)

percfilled <- total_sites %>% 
  mutate(percent_filled = 100*(total_trees/all_sites)) %>% 
  select(CSA2020, percent_filled)

percfilled %>% 
  ggplot(aes(percent_filled) ) +
  geom_density()

total_sites %>% summary()

#histogram
ggplot(total_sites, aes(all_sites)) +
  geom_histogram()

ggplot(total_sites, aes(total_trees)) +
  geom_histogram()

```


#9 Total Abundance by neighborhood
```{r}
#need to exclude unknown trees
NBabund <-no_vacant %>% 
  filter(Species != "unknown tree") %>% 
  group_by(CSA2020, Species) %>% 
  summarize(abund=sum(present)) %>% 
  select(CSA2020, Species, abund) 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)


```

#10 Richness and evenness
```{r}
richeven <- community_structure(NBabund, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020")


#norm by total trees
rich_tot <- richeven %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(rich_tot = richness/total_trees) %>% 
  select(CSA2020, rich_tot)
```

#11 Classify DBH (s/m/L)
```{r}
DBHclass <- no_vacant %>%
  mutate(size_class=ifelse(DBH>=0&DBH<=5, "small", ifelse(DBH>5&DBH<20, "medium", ifelse(DBH>=20, "large", 999)))) %>% 
  group_by(CSA2020, size_class) %>% 
  summarize(n=length(DBH)) %>% 
  pivot_wider(names_from = size_class, values_from = n, values_fill = 0) 

```

#12 % Filled with Large Trees
```{r}
perc_large <- total_sites %>% 
  left_join(DBHclass, by = "CSA2020") %>% 
  mutate(PercLarge_sites = (large/all_sites)*100) %>% 
  select(CSA2020, PercLarge_sites)

```

#13 % size
```{r}
PercSize <- DBHclass %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(PercSmall = small/total_trees *100) %>% 
  mutate(PercMedium = medium/total_trees*100) %>% 
  mutate(PercLarge = large/total_trees*100) %>% 
  select(CSA2020,PercSmall, PercMedium, PercLarge)
```

#14 Average DBH by neighborhood
```{r}
AvgDBH <- no_vacant %>% 
  select(CSA2020,DBH)%>% 
  group_by(CSA2020) %>% 
  summarise(AvgDBH = mean(DBH), .groups = 'drop')

#convert conditions to numeric 
```

#15 CV of all DBH measurements by neighborhood
```{r}
CvDBH<- no_vacant %>% 
  group_by(CSA2020) %>% 
  summarise(mean = mean(DBH), 
            sd = sd(DBH)) %>% 
  mutate(cv = sd/mean) %>% 
  select(cv, CSA2020)
```

#16 Total Small Trees
```{r}
Total_SM <- no_vacant %>% 
  filter(DBH>=0&DBH<=5) %>% 
  group_by(CSA2020) %>% 
  summarize(total_sm_trees =sum(present))

```

#17 SM Abund/Rich/evenness
```{r}
NBabundSM <-no_vacant %>% 
  filter(DBH>=0&DBH<=5) %>% 
  group_by(CSA2020, Species) %>% 
  summarize(abund=sum(present)) #%>% 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)

#SMabund_area <- NBabundSM %>% 
#   left_join(LandCover, by = "CSA2020") %>% 
#   mutate(SM_abund_area = abund/total_area_in_pixels) %>% 
#   select(SM_abund_area, CSA2020)

#calculate richness and eveness for small trees
richevenSM <- community_structure(NBabundSM, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020") %>% 
  rename(richness_SM = richness) %>% 
  rename(Evar_SM = Evar)


#norm total trees
sm_rich_tot <- richevenSM %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(sm_rich_tot = richness_SM/total_trees) %>% 
  select(sm_rich_tot, CSA2020)

```

#18 Total Large Trees
```{r}
Total_L <- no_vacant %>% 
  filter(DBH>20) %>% 
  group_by(CSA2020) %>% 
  summarize(total_L_trees =sum(present))

```

#19 L Abund/rich/even
```{r}
NBabundL <-no_vacant %>% 
  filter(DBH>=20) %>% 
  group_by(CSA2020, Species) %>% 
  summarize(abund=sum(present)) #%>% 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)

# #norm abundance by area 
# abund_area_L <- NBabundL %>% 
#   left_join(LandCover, by = "CSA2020") %>% 
#   mutate(abund_area_L = abund/total_area_in_pixels) %>% 
#   select(abund_area_L)

#calculate richness and evenness for Large trees
richevenL <- community_structure(NBabundL, abundance.var="abund", time.var= NULL, replicate.var = "CSA2020") %>% 
  rename(richness_L = richness) %>% 
  rename(Evar_L = Evar)




#richness norm by total trees
L_rich_tot <- richevenL %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(L_rich_tot = richness_L/total_trees) %>% 
  select(L_rich_tot, CSA2020)
```

#20 Native or introduced 
```{r}

species <- no_vacant %>%
  group_by(CSA2020) %>% 
  select(CSA2020, COMMON, Species) 

write.csv(species, file = '../../BaltimoreStreetTreeProject_Large_Data/tree_species.csv')

common_name <- no_vacant %>% 
  select(COMMON, Species) %>% 
  unique()

#adding column Nat = native, introduced, or N/I  
native <- species %>% 
    mutate(Nat = 
             #prunus serrulata mislabeled common name hackberry
             #INTRODUCED
             case_when(
               COMMON=='Acer truncatum'|COMMON=='Amur corktree'|COMMON=='Amur maackia'|
                 COMMON=='apple,common'|COMMON=='apricot'|COMMON=='arborvitae, Oriental'|
                 COMMON=='ash, black'|COMMON=='banana'|COMMON=='beech, European'|
                 COMMON=='birch, Asian white'|COMMON=='birch, Asian white'|COMMON=='birch, European white'|
                 COMMON=='birch, paper'|COMMON=='buckthorn, common'|COMMON=='Caucasian linden'|
                 COMMON=='cedar, Atlas'|COMMON=='cedar, deodar'|COMMON=='chastetree'|COMMON=='cherry, Higan'|
                 COMMON=='cherry, Japanese flowering'|COMMON=='cherry, Sargent'|COMMON=='cherry, sweet'|
                 COMMON=='cherry, Taiwan'|COMMON=='cherry, Yoshino'|COMMON=='cherrylaurel, Carolina'|
                 COMMON=='chestnut, Chinese'|COMMON=='crapemyrtle, common'|COMMON=="crapemyrtle, queen's"|
                 COMMON=='cypress, Italian'|COMMON=='cypress, Leyland'|COMMON=='dawn redwood'|
                 COMMON=='dogwood, corneliancherry'|COMMON=='dogwood, Kousa'|COMMON=='elm, Chinese'|
                 COMMON=='elm, rock'|COMMON=='elm, Siberian'|COMMON=='elm, slippery'|
                 COMMON=='elm, winged'|COMMON=='euonymus, spp.'|COMMON=='falsecypress, Hinoki'|
                 COMMON=='falsecypress, Japanese'|COMMON=='falsecypress, Lawson'|
                 COMMON=='falsecypress, Nootka'|COMMON=='falsecypress, spp.'|COMMON=='fig, common'|
                 COMMON=='filbert, Turkish'|COMMON=='fir, fraser'|COMMON=='fir, white'|
                 COMMON=='Fraxinus angustifolia'|COMMON=='ginkgo'|COMMON=='goldenraintree'|
                 COMMON=='hardy rubbertree'|COMMON=='hawthorn, Washington'|COMMON=='hickory, shellbark'|
                 COMMON=='holly, Chinese'|COMMON=='holly, English'|COMMON=="holly, Foster's"|
                 COMMON=='holly, Japanese'|COMMON=='holly, myrtle-leaf'|COMMON=='hornbeam, European'|
                 COMMON=='horsechestnut'|COMMON=='horsechestnut, red'|COMMON=='Japanese cryptomeria'|
                 COMMON=='Japanese pagodatree'|COMMON=='juniper, Chinese'|COMMON=='katsuratree'|
                 COMMON=='Kentucky coffeetree'|COMMON=='larch, American'|COMMON=='larch, European'|
                 COMMON=='lilac, Chinese tree'|COMMON=='lilac, common'|COMMON=='lilac, Japanese tree'|
                 COMMON=='linden, littleleaf'|COMMON=='linden, silver'|COMMON=='Liquidambar formosana'|
                 COMMON=='magnolia, saucer'|COMMON=='magnolia, southern'|COMMON=='magnolia, star'|
                 COMMON=='maple, Amur'|COMMON=='maple, black'|COMMON=='maple, Freeman'|COMMON=='maple, hedge'|
                 COMMON=='maple, Japanese'|COMMON=='maple, Miyabe'|COMMON=='maple, Norway'|
                 COMMON=='maple, paperbark'|COMMON=='maple, sycamore'|COMMON=='maple, Tatarian'|
                 COMMON=='maple, trident'|COMMON=='mimosa'|COMMON=='mountainash, European'|
                 COMMON=='mulberry, paper'|COMMON=='mulberry, red'|COMMON=='mulberry, white'|
                 COMMON=='oak, bur'|COMMON=='oak, English'|COMMON=='oak, live'|COMMON=='oak, northern pin'|
                 COMMON=='oak, overcup'|COMMON=='oak, Shumard'|COMMON=='oak, swamp chestnut'|
                 COMMON=='oak, Texas red'|COMMON=='oak, water'|COMMON=='oak, willow'|COMMON=='Oak, Willow'|
                 COMMON=='oriental raisin tree'|COMMON=='osage-orange'|COMMON=='peach, common'|
                 COMMON=='pear, callery'|COMMON=='pear, common'|COMMON=='pecan'|COMMON=='persian parrotia'|
                 COMMON=='photinia, Chinese'|COMMON=='photinia, fraser'|COMMON=='pine, Austrian'|
                 COMMON=='pine, Japanese black'|COMMON=='pine, Japanese white'|COMMON=='pine, mugo'|
                 COMMON=='pine, longleaf'|COMMON=='pine, red'|COMMON=='pistache, Chinese'|
                 COMMON=='planetree, London'|COMMON=='plum, cherry'|COMMON=='poplar, balsam'|
                 COMMON=='poplar, Lombardy'|COMMON=='poplar, white'|COMMON=='privet, glossy'|
                 COMMON=='rose-of-sharon'|COMMON=='royal paulownia'|COMMON=='smoketree, American'|
                 COMMON=='snowbell, American'|COMMON=='snowbell, Japanese'|COMMON=='Sorbus alnifolia'|
                 COMMON=='spruce, Colorado'|COMMON=='spruce, Norway'|COMMON=='spruce, Oriental'|
                 COMMON=='spruce, Serbian'|COMMON=='spruce, white'|COMMON=='sugarberry'|
                 COMMON=='sweet gum, Chinese'|COMMON=='sweetgum, Oriental'|COMMON=='Syringa pekinensis'|
                 COMMON=='Tilia mongolica'|COMMON=='Tilia x euchlora'|COMMON=='tree of heaven'|
                 COMMON=='viburnum, rusty blackhaw'|COMMON=='viburnum, spp.'|COMMON=='walnut, English'|
                 COMMON=='waxmyrtle, Southern'|COMMON=='willow, corkscrew'|COMMON=='willow, pussy'|
                 COMMON=='willow, weeping'|COMMON=='yellowwood'|COMMON=='zelkova, Chinese'|
                 COMMON=='zelkova, Japanese'|COMMON=='linden, American'|Species=='Poncirus trifoliata'|
                 Species=='Malus pumila'|Species=='Pinus sylvestris'|Species=='Pseudotsuga menziesii'|
                 Species=='Catalpa speciosa'|Species=='Catalpa bignon'|Species =='Fraxinus excelsior'|
                 COMMON=='catalpa, southern'|COMMON=='baldcypress, common'|Species=='Syringa pekinensis'|
                 Species=='Thuja plicata'|Species=='Picea abies'|Species=='Zelkova serrata'|
                 Species=='Acer truncatum'|Species=='Tilia mongolica'|Species=='persimmon, Japanese'
               ~'introduced',
              #NATIVE
              COMMON=='arborvitae spp.'|COMMON=='Aesculus parviflora'|COMMON=='arborvitae, eastern'|
                COMMON=='ash, black'|COMMON=='ash, green'|COMMON=='ash, pumpkin'|
                COMMON=='ash, white'|COMMON=='aspen, bigtooth'|COMMON=='beech, American'|
                COMMON=='birch, gray'|COMMON=='birch, river'|COMMON=='blackgum'|COMMON=='Blackgum'|
                COMMON=='boxelder'|COMMON=='buckeye, Ohio'|COMMON=='buckeye, red'|COMMON=='buckeye, yellow'|
                COMMON=='butternut'|COMMON == 'cherry, black'|COMMON=='chestnut, American'|
                COMMON=='chokecherry, common'|COMMON=='cottonwood, eastern'|COMMON=='crabapple, southern'|
                COMMON=='cypress, Arizona'|COMMON=='dogwood, flowering'|COMMON=='elm, American'|
                COMMON=='elm, smoothleaf'|COMMON=='filbert, American'|COMMON=='fringetree, white'|
                COMMON=='hackberry, common'|COMMON=='hawthorn, cockspur'|COMMON=='hawthorn, green'|
                COMMON=='hemlock, eastern'|COMMON=='hickory, bitternut'|COMMON=='hickory, mockernut'|
                COMMON=='hickory, pignut'|COMMON=='hickory, shagbark'|COMMON=='honeylocust'|
                COMMON=='honeylocust, thornless'|COMMON=='honeysuckle, spp.'|COMMON=='hophornbeam, American'|
                COMMON=='hornbeam, American'|COMMON=='inkberry'|COMMON=='locust, black'|
                COMMON=='magnolia, sweetbay'|COMMON=='maple, red'|COMMON=='maple, silver'|
                COMMON=='maple, sugar'|COMMON=='nannyberry'|COMMON=='ninebark, common'|COMMON=='oak, black'|
                COMMON=='Oak, Black'|COMMON=='oak, blackjack'|COMMON=='oak, chestnut'|
                COMMON=='oak, chinkapin'|COMMON=='oak, northern red'|COMMON=='oak, pin'|COMMON=='oak, post'|
                COMMON=='oak, sawtooth'|COMMON=='oak, scarlet'|COMMON=='oak, shingle'|
                COMMON=='oak, southern red'|COMMON=='oak, swamp white'|COMMON=='Oak, swamp white'|
                COMMON=='oak, white'|COMMON=='pawpaw, common'|COMMON=='pine, pond'|COMMON=='pine, shortleaf'|
                COMMON=='pine, Virginia'|COMMON=='plum, American'|COMMON=='redbud, eastern'|
                COMMON=='redcedar, eastern'|COMMON=='sassafras'|COMMON=='serviceberry, Allegheny'|
                COMMON=='serviceberry, downy'|COMMON=='serviceberry, shadblow'|COMMON=='serviceberry, spp.'|
                COMMON=='sumac, staghorn'|COMMON=='sweetgum, American'|COMMON=='sycamore, American'|
                COMMON=='tuliptree'|COMMON=='walnut, black'|COMMON=='willow, black'|
                COMMON=='witchhazel, common'|Species == 'Pinus strobus'|Species == 'Diospyros virginiana'|
                Species=='Ilex opaca'|Species=='Crataegus mollis'|COMMON=='pine, loblolly'|
                Species=='Magnolia acuminata'|Species=='Ulmus americana'|Species=='Acer saccharum'|
                COMMON=='elm, hybrid'
              ~'native',
              #N/I = AT GENUS LEVEL (GENUS CONTAINS BOTH NATIVE AND INTRODUCED SPECIES)
              COMMON=='cherry/plum, spp.'|COMMON=='crabapple, flowering'|COMMON=='hawthorn, spp.'|COMMON=='yew, spp.'|Species=='Prunus spp.'  ~ 'N/I',
              #double check that all species were covered
              TRUE ~ 'not done yet'))

#calculate %native - single value for each neighborhood
  
PercNative <- native %>%
  group_by(CSA2020, Nat) %>% 
  summarize(n=length(Nat)) %>%
  pivot_wider(names_from = Nat, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal) %>% 
  mutate(PercNat = (native/total_trees)*100) %>% 
  select(CSA2020, PercNat)

```

#21 Species Abundance per NB
```{r}
species_abund <- species %>%
  group_by(CSA2020, Species) %>% 
  summarize(n=length(Species)) %>%
  pivot_wider(names_from = Species, values_from = n, values_fill = 0) 

write_csv(species_abund,'../input_data/Species_Abund.csv')
```

#22 Pit Area Calculation
```{r}

#this is at the tree level, our other calculations are at the nb level
pit_size <- clean %>% 
  mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         , pit_area   = length*width
         ) %>% 
  select(CSA2020, pit_area, X_COORD, Y_COORD, DBH)

ggplot(pit_size, aes(x=pit_area)) + geom_histogram()

cor.test(pit_size$pit_area, pit_size$DBH)
# r=0.342

large_pit <- pit_size %>% 
  mutate(size_threshhold = ifelse(pit_area >= 32, "great32", "less32" )) %>% 
  group_by(CSA2020, size_threshhold) %>% 
  summarize(n=length(size_threshhold)) %>% 
  pivot_wider(names_from = size_threshhold, values_from = n, values_fill = 0) %>% 
  left_join(total_sites, by = "CSA2020") %>% 
  mutate(Perc32 = great32/all_sites*100) %>% 
  select(CSA2020, Perc32)


```

#23 Tree Condition
```{r}
#sum for each condition in a neighborhood
#pivot
condition <- no_vacant %>%
  group_by(CSA2020, CONDITION) %>% 
  summarize(n=length(CONDITION)) %>% 
  pivot_wider(names_from = CONDITION, values_from = n, values_fill = 0)

# numeric condition
condition_grade <- no_vacant %>% 
  filter(CONDITION!="Sprout"&CONDITION!="Sprouts") %>% 
  mutate(grade = 
  case_when(
    CONDITION == "Poor" ~ 1,
    CONDITION == "Fair" ~ 2,
    CONDITION == "Good" ~ 3,
    )) %>% 
  select(CSA2020,grade) 


#Condition vs pit size
#%>% 
  # left_join(no_vacant_pit, by = "CSA2020")
```

#24 Condition and Pit Area
```{r}

no_vacant |> 
  filter(CONDITION!="Sprout"&CONDITION!="Sprouts") |>
   mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         , pit_area   = length*width
         ) |>
  ggplot(aes(CONDITION, pit_area)) +
  geom_boxplot()
  
```

#23 % good/poor trees
```{r}
perc_condition <- NBTotal %>% 
  left_join(condition, by = "CSA2020") %>% 
  mutate(PercGood = 100*(Good/total_trees)) %>% 
  mutate(PercPoor = 100*(Poor/total_trees)) %>% 
  select(CSA2020, PercPoor, PercGood)
```

#24 Flowering Trees
```{r}
# read in flowering data
flowering <- read_csv("../../BaltimoreStreetTreeProject_Large_Data/Flowering Data.csv", show_col_types = FALSE)

flower_y_n <- flowering %>% 
  select(CSA2020, Flowering) %>% 
  filter(Flowering =="yes" | Flowering == "no")

showiness <- flowering %>% 
  select(CSA2020, Visibility)

PercFlower <- flower_y_n %>%
  group_by(CSA2020, Flowering) %>% 
  summarize(n=length(Flowering)) %>%
  pivot_wider(names_from = Flowering, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  mutate(PercFlor = (yes/total_trees)*100) %>% 
  select(CSA2020, PercFlor)

PercShowy <- showiness %>% 
  group_by(CSA2020, Visibility) %>% 
  summarize(n=length(Visibility)) %>%
  pivot_wider(names_from = Visibility, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal, by = "CSA2020") %>% 
  rename("showy" = '3') %>% 
  mutate(PercShowy = (showy/total_trees)*100) %>% 
  select(CSA2020, PercShowy)
```

#24 5 Hottest NB Top 5 Species 
```{r}
small_trees_abund <- no_vacant %>%
  filter(DBH>=0&DBH<=5) %>%  
  group_by(CSA2020, Species) %>% 
  summarize(n=length(Species)) %>%
  pivot_wider(names_from = Species, values_from = n, values_fill = 0)

small_abund_temp <- NBtemperature %>%
  filter(CSA2020 != "Dickeyville/Franklintown" & CSA2020 != "Unassigned -- Jail") %>% 
  left_join(small_trees_abund, by = "CSA2020") %>% 
  arrange(desc(avg_temp)) %>% 
  select(-avg_temp)

#top species for Madison
madison_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Madison/East End") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(madison_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Madison/East End"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Patterson Park
patterson_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Patterson Park North & East") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(patterson_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Patterson Park North & East"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# oldtown
oldtown_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Oldtown/Middle East") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(oldtown_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Oldtown/Middle East"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Highlandtown
highland_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Highlandtown") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(highland_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Highlandtown"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Downtown/Seton Hill
downtown_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Downtown/Seton Hill") %>% 
  gather(key = "TreeSpecies", value = "Count", -CSA2020) %>%
  slice_max(order_by = Count, n = 5)

ggplot(downtown_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Downtown/Seton Hill"), x = "TreeSpecies", y = "Count") +

  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#25 5 Coolest NB Top 5 Species
```{r}
#mount washington
mtwash_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Mount Washington/Coldspring") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(mtwash_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Mount Washington/Coldspring"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#forest park
forpark_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Forest Park/Walbrook") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(forpark_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Forest Park/Walbrook"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#edmonson village
edmonson_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Edmonson Village") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(edmonson_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Edmonson Village"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#greater roland park
roland_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Greater Roland Park/Poplar Hill") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(roland_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Greater Roland Park/Poplar Hill"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#beechfield
beechfield_top_species <- small_abund_temp %>% 
  filter(CSA2020 == "Beechfield/Ten Hills/West Hills") %>% #filter for neighborhood
  gather(key = "TreeSpecies", value = "Count", -CSA2020)  %>% # make long
  slice_max(order_by = Count, n = 5) #slice top 5 species

ggplot(beechfield_top_species, aes(x = TreeSpecies, y = Count, fill = TreeSpecies)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Top Tree Species in", "Beechfield/Ten Hills/West Hills"), x = "TreeSpecies", y = "Count") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set1") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

#25 Combine all dependent variables
```{r}
 DepVar <- percfilled %>% 
  left_join(total_sites) %>% 
  left_join(richeven, by = "CSA2020") %>% 
  left_join(rich_tot, by = "CSA2020") %>% 
  left_join(AvgDBH, by = "CSA2020") %>% 
  
  
  left_join(PercNative, by = "CSA2020") %>% 
  
  left_join(perc_condition, by = "CSA2020") %>% 
  left_join(large_pit, by = "CSA2020") %>% 
  
  left_join(PercFlower, by = "CSA2020") %>% 
  left_join(PercShowy, by = "CSA2020")


  
write_csv(DepVar,'../input_data/Dependent_variable.csv')  
```




