---
title: "Tree data"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---

This file contains all dependent variable data for analysis

# 0 set up 
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 
library(dplyr)
library(vegan)
library(codyn)

getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```

# 1 read in street tree data 
```{r include=FALSE}
street_tree <- 
  st_read('../../BaltimoreStreetTreeProject_Large_Data\\bc_forestry_trees_20190319\\bc_forestry_trees_20190319.shp'
, as_tibble = TRUE) %>% 
  filter(LOC_TYPE == "Street") 
```

# 2 combine with BNIA neighborhood boundaries
```{r include=FALSE}
bmore_bnia <- st_read('../../BaltimoreStreetTreeProject_Large_Data\\Community_Statistical_Areas_(CSAs)__Reference_Boundaries/Community_Statistical_Areas_(CSAs)__Reference_Boundaries.shp') |> #BNIA community delineations
  st_make_valid() |>
  st_transform(crs = st_crs(street_tree))

combined <-  
  street_tree %>%
  st_join(bmore_bnia) %>%
  st_drop_geometry()

combined |> 
  write_csv(paste0('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_', Sys.Date(), '.csv'))
```

#3 Read in street tree + neighborhood file
```{r include=FALSE}
combined <-
  read_csv('../../BaltimoreStreetTreeProject_Large_Data/street_trees_with_neigh_attributes_2023-07-27.csv')
```

#4 Clean combined data
```{r}
clean <- combined %>% 
  select(-ID,-OnStr, -OBSERV2, -OBSERV3, -HARD_SCAPE, -PARCELID, -COLLECTOR, -Inv_Time, -INSPECT_DT, -Inspect_TM, -Notes, -Link, -UniqueID, -UTILITIES, -Neigh) %>% 
  rename(CSA2010 = CSA2020) #doing this because the bnia boundaries only included from 2020 but all demographic data had 2010 boundaries, but they're the same I checked
```

#5 Filter tree data for only present tree species
```{r include=FALSE}
no_vacant <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species!='Vacant Site'&Species!='Vacant Potential'&Species!='Stump'&Species!='Vacant Site Not Suitable'&Species!='NA'&Species!='unknown shrub'&Species!="Z Add 01"&Species!=" "&Species!="Dead")%>%
  filter(CONDITION!="Dead"&CONDITION!="Stump"&CONDITION!="Sprout"&CONDITION!="Absent")%>% 
  mutate(present=1) #for each tree make a present column with 1

```

#6 Total Potential Sites
```{r}
vacant_potential <- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species=='Vacant Site'|Species=='Vacant Potential'|Species=='Stump'|Species=='NA'|Species=='unknown shrub'|Species=="Z Add 01"|Species=="Dead")%>% 
  mutate(vacant_present=1) %>%   #for each vacant potential site present column with 1 
  group_by(CSA2010) %>% 
  summarize(sum_vacant_present =sum(vacant_present)) 


  
```

#7 Total # Trees per Neighborhood - how filled is a neighborhood
```{r}
NBTotal <-no_vacant %>% 
  group_by(CSA2010) %>% 
  summarize(total_trees =sum(present)) %>% 
  select(total_trees, CSA2010)

total_sites <- vacant_potential %>% 
  left_join(NBTotal, by = "CSA2010") %>% 
  mutate(all_sites = sum_vacant_present + total_trees, 
         percent_filled = 100*(total_trees/all_sites))

total_sites %>% 
  ggplot(aes(percent_filled) ) +
  geom_density()

total_sites %>% summary()

total_sites %>% 
  ggplot(aes(total_trees, percent_filled)) +
  geom_point() +
  NULL

#normalize by total sites  

```

#8 Total Abundance by neighborhood
```{r}
#need to exclude unknown trees
NBabund <-no_vacant %>% 
  filter(Species != "unknown tree") %>% 
  group_by(CSA2010, Species) %>% 
  summarize(abund=sum(present)) %>% 
  select(CSA2010, Species, abund) 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)


```

#9 Richness and evenness
```{r}
richeven <- community_structure(NBabund, abundance.var="abund", time.var= NULL, replicate.var = "CSA2010")


#norm by total trees
rich_tot <- richeven %>% 
  left_join(NBTotal, by = "CSA2010") %>% 
  mutate(rich_tot = richness/total_trees)

#norm by total sites

```

#10 Classify DBH (s/m/L)
```{r}
DBHclass <- no_vacant %>%
  mutate(size_class=ifelse(DBH>=0&DBH<=5, "small", ifelse(DBH>5&DBH<20, "medium", ifelse(DBH>=20, "large", 999)))) %>% 
  group_by(CSA2010, size_class) %>% 
  summarize(n=length(DBH)) %>% 
  pivot_wider(names_from = size_class, values_from = n, values_fill = 0)

```

#11 Average DBH by neighborhood
```{r}
AvgDBH <- no_vacant %>% 
  select(CSA2010,DBH)%>% 
  group_by(CSA2010) %>% 
  summarise(AvgDBH = mean(DBH), .groups = 'drop')
```

#12 CV of all DBH measurements by neighborhood
```{r}
CvDBH<- no_vacant %>% 
  group_by(CSA2010) %>% 
  summarise(mean = mean(DBH), 
            sd = sd(DBH)) %>% 
  mutate(cv = sd/mean) %>% 
  select(cv, CSA2010)
```

#13 Total Small Trees
```{r}
Total_SM <- no_vacant %>% 
  filter(DBH>=0&DBH<=5) %>% 
  group_by(CSA2010) %>% 
  summarize(total_sm_trees =sum(present))

```

#14 SM Abund/Rich/evenness
```{r}
NBabundSM <-no_vacant %>% 
  filter(DBH>=0&DBH<=5) %>% 
  group_by(CSA2010, Species) %>% 
  summarize(abund=sum(present)) #%>% 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)

#SMabund_area <- NBabundSM %>% 
#   left_join(LandCover, by = "CSA2010") %>% 
#   mutate(SM_abund_area = abund/total_area_in_pixels) %>% 
#   select(SM_abund_area, CSA2010)

#calculate richness and eveness for small trees
richevenSM <- community_structure(NBabundSM, abundance.var="abund", time.var= NULL, replicate.var = "CSA2010") %>% 
  rename(richness_SM = richness) %>% 
  rename(Evar_SM = Evar)


#norm total trees
sm_rich_tot <- richevenSM %>% 
  left_join(NBTotal, by = "CSA2010") %>% 
  mutate(sm_rich_tot = richness_SM/total_trees) %>% 
  select(sm_rich_tot, CSA2010)

```

#16 Total Large Trees
```{r}
Total_L <- no_vacant %>% 
  filter(DBH>20) %>% 
  group_by(CSA2010) %>% 
  summarize(total_L_trees =sum(present))

```

#17 L Abund/rich/even
```{r}
NBabundL <-no_vacant %>% 
  filter(DBH>=20) %>% 
  group_by(CSA2010, Species) %>% 
  summarize(abund=sum(present)) #%>% 
#  pivot_wider(names_from = Species, values_from = abund, values_fill = 0)

# #norm abundance by area 
# abund_area_L <- NBabundL %>% 
#   left_join(LandCover, by = "CSA2010") %>% 
#   mutate(abund_area_L = abund/total_area_in_pixels) %>% 
#   select(abund_area_L)

#calculate richness and evenness for Large trees
richevenL <- community_structure(NBabundL, abundance.var="abund", time.var= NULL, replicate.var = "CSA2010") %>% 
  rename(richness_L = richness) %>% 
  rename(Evar_L = Evar)




#richness norm by total trees
L_rich_tot <- richevenL %>% 
  left_join(NBTotal, by = "CSA2010") %>% 
  mutate(L_rich_tot = richness_L/total_trees) %>% 
  select(L_rich_tot, CSA2010)
```

#18 Native or introduced 
```{r}

species <- no_vacant %>%
  group_by(CSA2010) %>% 
  select(CSA2010, COMMON, Species) 

common_name <- no_vacant %>% 
  select(COMMON, Species) %>% 
  unique()

#adding column Nat = native, introduced, or N/I  
native <- species %>% 
    mutate(Nat = 
             #prunus serrulata mislabeled common name hackberry
             #INTRODUCED
             case_when(
               COMMON=='Acer truncatum'|COMMON=='Amur corktree'|COMMON=='Amur maackia'|
                 COMMON=='apple,common'|COMMON=='apricot'|COMMON=='arborvitae, Oriental'|
                 COMMON=='ash, black'|COMMON=='banana'|COMMON=='beech, European'|
                 COMMON=='birch, Asian white'|COMMON=='birch, Asian white'|COMMON=='birch, European white'|
                 COMMON=='birch, paper'|COMMON=='buckthorn, common'|COMMON=='Caucasian linden'|
                 COMMON=='cedar, Atlas'|COMMON=='cedar, deodar'|COMMON=='chastetree'|COMMON=='cherry, Higan'|
                 COMMON=='cherry, Japanese flowering'|COMMON=='cherry, Sargent'|COMMON=='cherry, sweet'|
                 COMMON=='cherry, Taiwan'|COMMON=='cherry, Yoshino'|COMMON=='cherrylaurel, Carolina'|
                 COMMON=='chestnut, Chinese'|COMMON=='crapemyrtle, common'|COMMON=="crapemyrtle, queen's"|
                 COMMON=='cypress, Italian'|COMMON=='cypress, Leyland'|COMMON=='dawn redwood'|
                 COMMON=='dogwood, corneliancherry'|COMMON=='dogwood, Kousa'|COMMON=='elm, Chinese'|
                 COMMON=='elm, rock'|COMMON=='elm, Siberian'|COMMON=='elm, slippery'|
                 COMMON=='elm, winged'|COMMON=='euonymus, spp.'|COMMON=='falsecypress, Hinoki'|
                 COMMON=='falsecypress, Japanese'|COMMON=='falsecypress, Lawson'|
                 COMMON=='falsecypress, Nootka'|COMMON=='falsecypress, spp.'|COMMON=='fig, common'|
                 COMMON=='filbert, Turkish'|COMMON=='fir, fraser'|COMMON=='fir, white'|
                 COMMON=='Fraxinus angustifolia'|COMMON=='ginkgo'|COMMON=='goldenraintree'|
                 COMMON=='hardy rubbertree'|COMMON=='hawthorn, Washington'|COMMON=='hickory, shellbark'|
                 COMMON=='holly, Chinese'|COMMON=='holly, English'|COMMON=="holly, Foster's"|
                 COMMON=='holly, Japanese'|COMMON=='holly, myrtle-leaf'|COMMON=='hornbeam, European'|
                 COMMON=='horsechestnut'|COMMON=='horsechestnut, red'|COMMON=='Japanese cryptomeria'|
                 COMMON=='Japanese pagodatree'|COMMON=='juniper, Chinese'|COMMON=='katsuratree'|
                 COMMON=='Kentucky coffeetree'|COMMON=='larch, American'|COMMON=='larch, European'|
                 COMMON=='lilac, Chinese tree'|COMMON=='lilac, common'|COMMON=='lilac, Japanese tree'|
                 COMMON=='linden, littleleaf'|COMMON=='linden, silver'|COMMON=='Liquidambar formosana'|
                 COMMON=='magnolia, saucer'|COMMON=='magnolia, southern'|COMMON=='magnolia, star'|
                 COMMON=='maple, Amur'|COMMON=='maple, black'|COMMON=='maple, Freeman'|COMMON=='maple, hedge'|
                 COMMON=='maple, Japanese'|COMMON=='maple, Miyabe'|COMMON=='maple, Norway'|
                 COMMON=='maple, paperbark'|COMMON=='maple, sycamore'|COMMON=='maple, Tatarian'|
                 COMMON=='maple, trident'|COMMON=='mimosa'|COMMON=='mountainash, European'|
                 COMMON=='mulberry, paper'|COMMON=='mulberry, red'|COMMON=='mulberry, white'|
                 COMMON=='oak, bur'|COMMON=='oak, English'|COMMON=='oak, live'|COMMON=='oak, northern pin'|
                 COMMON=='oak, overcup'|COMMON=='oak, Shumard'|COMMON=='oak, swamp chestnut'|
                 COMMON=='oak, Texas red'|COMMON=='oak, water'|COMMON=='oak, willow'|COMMON=='Oak, Willow'|
                 COMMON=='oriental raisin tree'|COMMON=='osage-orange'|COMMON=='peach, common'|
                 COMMON=='pear, callery'|COMMON=='pear, common'|COMMON=='pecan'|COMMON=='persian parrotia'|
                 COMMON=='photinia, Chinese'|COMMON=='photinia, fraser'|COMMON=='pine, Austrian'|
                 COMMON=='pine, Japanese black'|COMMON=='pine, Japanese white'|COMMON=='pine, mugo'|
                 COMMON=='pine, longleaf'|COMMON=='pine, red'|COMMON=='pistache, Chinese'|
                 COMMON=='planetree, London'|COMMON=='plum, cherry'|COMMON=='poplar, balsam'|
                 COMMON=='poplar, Lombardy'|COMMON=='poplar, white'|COMMON=='privet, glossy'|
                 COMMON=='rose-of-sharon'|COMMON=='royal paulownia'|COMMON=='smoketree, American'|
                 COMMON=='snowbell, American'|COMMON=='snowbell, Japanese'|COMMON=='Sorbus alnifolia'|
                 COMMON=='spruce, Colorado'|COMMON=='spruce, Norway'|COMMON=='spruce, Oriental'|
                 COMMON=='spruce, Serbian'|COMMON=='spruce, white'|COMMON=='sugarberry'|
                 COMMON=='sweet gum, Chinese'|COMMON=='sweetgum, Oriental'|COMMON=='Syringa pekinensis'|
                 COMMON=='Tilia mongolica'|COMMON=='Tilia x euchlora'|COMMON=='tree of heaven'|
                 COMMON=='viburnum, rusty blackhaw'|COMMON=='viburnum, spp.'|COMMON=='walnut, English'|
                 COMMON=='waxmyrtle, Southern'|COMMON=='willow, corkscrew'|COMMON=='willow, pussy'|
                 COMMON=='willow, weeping'|COMMON=='yellowwood'|COMMON=='zelkova, Chinese'|
                 COMMON=='zelkova, Japanese'|COMMON=='linden, American'|Species=='Poncirus trifoliata'|
                 Species=='Malus pumila'|Species=='Pinus sylvestris'|Species=='Pseudotsuga menziesii'|
                 Species=='Catalpa speciosa'|Species=='Catalpa bignon'|Species =='Fraxinus excelsior'|
                 COMMON=='catalpa, southern'|COMMON=='baldcypress, common'|Species=='Syringa pekinensis'|
                 Species=='Thuja plicata'|Species=='Picea abies'|Species=='Zelkova serrata'|
                 Species=='Acer truncatum'|Species=='Tilia mongolica'|Species=='persimmon, Japanese'
               ~'introduced',
              #NATIVE
              COMMON=='arborvitae spp.'|COMMON=='Aesculus parviflora'|COMMON=='arborvitae, eastern'|
                COMMON=='ash, black'|COMMON=='ash, green'|COMMON=='ash, pumpkin'|
                COMMON=='ash, white'|COMMON=='aspen, bigtooth'|COMMON=='beech, American'|
                COMMON=='birch, gray'|COMMON=='birch, river'|COMMON=='blackgum'|COMMON=='Blackgum'|
                COMMON=='boxelder'|COMMON=='buckeye, Ohio'|COMMON=='buckeye, red'|COMMON=='buckeye, yellow'|
                COMMON=='butternut'|COMMON == 'cherry, black'|COMMON=='chestnut, American'|
                COMMON=='chokecherry, common'|COMMON=='cottonwood, eastern'|COMMON=='crabapple, southern'|
                COMMON=='cypress, Arizona'|COMMON=='dogwood, flowering'|COMMON=='elm, American'|
                COMMON=='elm, smoothleaf'|COMMON=='filbert, American'|COMMON=='fringetree, white'|
                COMMON=='hackberry, common'|COMMON=='hawthorn, cockspur'|COMMON=='hawthorn, green'|
                COMMON=='hemlock, eastern'|COMMON=='hickory, bitternut'|COMMON=='hickory, mockernut'|
                COMMON=='hickory, pignut'|COMMON=='hickory, shagbark'|COMMON=='honeylocust'|
                COMMON=='honeylocust, thornless'|COMMON=='honeysuckle, spp.'|COMMON=='hophornbeam, American'|
                COMMON=='hornbeam, American'|COMMON=='inkberry'|COMMON=='locust, black'|
                COMMON=='magnolia, sweetbay'|COMMON=='maple, red'|COMMON=='maple, silver'|
                COMMON=='maple, sugar'|COMMON=='nannyberry'|COMMON=='ninebark, common'|COMMON=='oak, black'|
                COMMON=='Oak, Black'|COMMON=='oak, blackjack'|COMMON=='oak, chestnut'|
                COMMON=='oak, chinkapin'|COMMON=='oak, northern red'|COMMON=='oak, pin'|COMMON=='oak, post'|
                COMMON=='oak, sawtooth'|COMMON=='oak, scarlet'|COMMON=='oak, shingle'|
                COMMON=='oak, southern red'|COMMON=='oak, swamp white'|COMMON=='Oak, swamp white'|
                COMMON=='oak, white'|COMMON=='pawpaw, common'|COMMON=='pine, pond'|COMMON=='pine, shortleaf'|
                COMMON=='pine, Virginia'|COMMON=='plum, American'|COMMON=='redbud, eastern'|
                COMMON=='redcedar, eastern'|COMMON=='sassafras'|COMMON=='serviceberry, Allegheny'|
                COMMON=='serviceberry, downy'|COMMON=='serviceberry, shadblow'|COMMON=='serviceberry, spp.'|
                COMMON=='sumac, staghorn'|COMMON=='sweetgum, American'|COMMON=='sycamore, American'|
                COMMON=='tuliptree'|COMMON=='walnut, black'|COMMON=='willow, black'|
                COMMON=='witchhazel, common'|Species == 'Pinus strobus'|Species == 'Diospyros virginiana'|
                Species=='Ilex opaca'|Species=='Crataegus mollis'|COMMON=='pine, loblolly'|
                Species=='Magnolia acuminata'|Species=='Ulmus americana'|Species=='Acer saccharum'|
                COMMON=='elm, hybrid'
              ~'native',
              #N/I = AT GENUS LEVEL (GENUS CONTAINS BOTH NATIVE AND INTRODUCED SPECIES)
              COMMON=='cherry/plum, spp.'|COMMON=='crabapple, flowering'|COMMON=='hawthorn, spp.'|COMMON=='yew, spp.'|Species=='Prunus spp.'  ~ 'N/I',
              #double check that all species were covered
              TRUE ~ 'not done yet'))

#calculate %native - single value for each neighborhood
  
PercNative <- native %>%
  group_by(CSA2010, Nat) %>% 
  summarize(n=length(Nat)) %>%
  pivot_wider(names_from = Nat, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal) %>% 
  mutate(PercNat = (native/total_trees)*100)

```

#19 Species Abundance per NB
```{r}
species_abund <- species %>%
  group_by(CSA2010, Species) %>% 
  summarize(n=length(Species)) %>%
  pivot_wider(names_from = Species, values_from = n, values_fill = 0) %>% 
  left_join(NBTotal)
```

#20 Pit size
```{r}
pit_size <- clean %>% 
  mutate(  length = as.numeric(ifelse(SPCELENGTH == '>20', 21, SPCELENGTH))
         , width  = as.numeric(ifelse(SPACEWIDTH == '>20', 21, SPACEWIDTH))
         , area   = length*width
         ) %>% 
  select(CSA2010, area)

```

#21 Tree Condition
```{r}
#sum for each condition in a neighborhood
#pivot

condition <- no_vacant %>%
  group_by(CSA2010, CONDITION) %>% 
  summarize(n=length(CONDITION)) %>% 
  pivot_wider(names_from = CONDITION, values_from = n, values_fill = 0)
```

#22 Combine all dependent variables
```{r}
 DepVar <- total_sites %>% 
  left_join(richeven, by = "CSA2010") %>% 
  left_join(AvgDBH, by = "CSA2010") %>% 
  left_join(CvDBH, by = "CSA2010") %>% 
  
  left_join(richevenSM, by = "CSA2010") %>% 
  left_join(sm_rich_tot, by = "CSA2010") %>% 
  
  left_join(richevenL, by = "CSA2010") %>%
  left_join(L_rich_tot, by = "CSA2010") %>% 
  
  left_join(vacant_potential, by = "CSA2010") %>%  
  
  
  left_join(PercNative, by = "CSA2010") %>% 
  left_join(vacant_potential, by = "CSA2010")  



  
write_csv(DepVar,'../input_data/Dependent_variable.csv')  
```




