---
title: "Analyses"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output:
  pdf_document:
    toc: yes
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: yes
    toc: yes
    toc_float: yes
    self_contained: yes
  word_document:
    toc: yes
---

This file will analyze street tree data at neighborhood level

# setup + loading packages
```{r include=FALSE}
# load packages

library(tidyverse)
library(sf) #simple features - vector GIS support
library(mapview) #quick web maps 
library(tidylog) #makes dplyr and tidyr verbose
library(dplyr)
library(vegan)
library(codyn)
library(ggplot2)

#set wd
# run baltimore neighborhoods to get clean file 
getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/') # means "up one level"
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE) # means "up one level"


```


#make clean
```{r include=FALSE}

 clean <- combined %>% 
select(-ID,-OnStr, -OBSERV2, -OBSERV3, -LOC_TYPE, -HARD_SCAPE, -PARCELID, -COLLECTOR, -Inv_Time, -INSPECT_DT, -Inspect_TM, -Notes, -Link) %>% 
  rename(Neighborhood = Community)
```

#filter tree data for only present tree species, create NB demographics data frame
```{r}
data<- clean %>% 
  mutate(Species=ifelse(SPP=='Acer spp.'|SPP=='Carya spp.'|SPP=='Cornus spp.'|SPP=='Cornus x'|SPP=='Ficus spp.'|SPP=='Fraxinus spp.'|SPP=='Hydrangea spp.'|SPP=='Ilex spp.'|SPP=='Ilex x'|SPP=='Juniperus spp.'|SPP=='Magnolia spp.'|SPP=='Magnolia x'|SPP=='Photinia spp.'|SPP=='Picea spp.'|SPP=='Populus spp.'|SPP=='Quercus spp.'|SPP=='Quercus x'|SPP=='Rhus spp.'|SPP=='Salix spp.'|SPP=='Ulmus spp.', "unknown tree", SPP))%>%
  filter(Species!='Vacant Site'&Species!='Vacant Potential'&Species!='Stump'&Species!='Vacant Site Not Suitable'&Species!='NA'&Species!='unknown shrub'&Species!="Z Add 01"&Species!=" "&Species!="Dead")%>%
  filter(CONDITION!="Dead"&CONDITION!="Stump"&CONDITION!="Sprout")%>% 
  mutate(present=1) %>% #for each tree make a present column with 1
  filter(Species!= "unknown tree") 

# OLD neighborhood demographics data frame
# NBdem<- data%>%
#   mutate(PercentWhite = White/Population*100) %>% #converting race data into percentages
#   mutate(PercentBlk = Blk_AfAm/Population*100) %>%
#   mutate(PercentNtv = AmInd_AkNa/Population*100) %>%
#   mutate(PercentAsian = Asian/Population*100) %>%
#   mutate(PercentNatHaw_Pac = NatHaw_Pac/Population*100) %>%
#   mutate(PercentLat_Hisp = Hisp_Lat/Population*100) %>%
#   mutate(PercentOther_Race = Other_Race/Population*100) %>%
#   select(Neighborhood, Population, PercentWhite, PercentBlk, PercentNtv, PercentAsian, PercentNatHaw_Pac, PercentLat_Hisp, PercentOther_Race) %>% 
#   group_by(Neighborhood) %>%
#  unique() 
#For each NB sum the number of trees by species (abundance)
```



```{r}
NBabund <-data %>% 
  group_by(Neighborhood, Species) %>% 
  summarize(abund=sum(present))

#NBabund|> 
#  write_csv(paste0('../../BaltimoreStreetTreeProject_Large_Data/trees_per_neighborhood_', Sys.Date(), '.csv')) # file of trees per neighborhood for Meghan
```

#calculate richness and evenness 
```{r}

richeven <-community_structure(NBabund, abundance.var="abund", time.var= NULL, replicate.var = "Neighborhood") %>%
 left_join(NBdem)
```
#vacant tree sites
```{r}
#filter for vacant sites by neighborhood

Vacant<-clean %>%
  filter(SPP=='Vacant Site'|SPP=='Vacant Potential'|SPP=='Stump'|SPP=='NA'|SPP=='unknown shrub'|SPP=='Z Add 01'|SPP==' '|SPP=='Dead') %>% #how would I group these together?
  mutate(present = 1) %>% 
  select(Neighborhood, SPP, present) %>% 
  group_by(Neighborhood) %>% 
  unique()
#number of vacant sites per neighborhood
NBvacant <-Vacant %>% 
  group_by(Neighborhood) %>% 
  summarize(vacant=sum(present))  
```

# some quick correlation analyses
```{r}
#percent white vs richness
cor.test(richeven$PercentWhite, richeven$richness)
plot(richeven$PercentWhite, richeven$richness)

#percent white vs evenness
cor.test(richeven$PercentWhite, richeven$Evar)
plot(richeven$PercentWhite, richeven$Evar)

#percent black richness
cor.test(richeven$PercentBlk, richeven$richness)
plot(richeven$PercentBlk, richeven$richness)

#percent black evenness
cor.test(richeven$PercentBlk, richeven$Evar)
plot(richeven$PercentBlk, richeven$Evar)

# #percent native american and alaska native
# cor.test(richeven$PercentNtv, richeven$richness)
# plot(richeven$PercentNtv, richeven$richness)

#percent Latino and Hispanic richness
cor.test(richeven$PercentLat_Hisp, richeven$richness)
plot(richeven$PercentLat_Hisp, richeven$richness)

#percent Latino and Hispanic evenness
cor.test(richeven$PercentLat_Hisp, richeven$Evar)
plot(richeven$PercentLat_Hisp, richeven$Evar)

#population vs richness
cor.test(richeven$Population, richeven$richness)
plot(richeven$Population, richeven$richness)
#population vs evenness
cor.test(richeven$Population, richeven$Evar)
plot(richeven$Population, richeven$Evar)
```

#DBH - classify trees by size
```{r}

DBHclass<-data%>%
  mutate(size_class=ifelse(DBH>=0&DBH<=5, "small", ifelse(DBH>5&DBH<20, "medium", ifelse(DBH>=20, "large", 999)))) %>% 
  group_by(Neighborhood, size_class) %>% 
  summarize(n=length(DBH)) %>% 
  pivot_wider(names_from = size_class, values_from = n, values_fill = 0)

#joining with neighborhood demographics
DBHdem <- DBHclass %>% 
  left_join(NBdem)

```

#average DBH by neighborhood
```{r}
NBDBHAvg <- data %>% 
  select(Neighborhood, DBH) %>% 
  group_by(Neighborhood) %>% 
  summarise(AvgDBH = mean(DBH), .groups = 'drop')

AvgDBHdem <- NBDBHAvg %>% 
  left_join(NBdem)

```
#correlations for DBH
```{r}
#percent black
cor.test(AvgDBHdem$PercentBlk,AvgDBHdem$AvgDBH)
plot(AvgDBHdem$PercentBlk,AvgDBHdem$AvgDBH)


cor.test(AvgDBHdem$PercentWhite,AvgDBHdem$AvgDBH)
plot(AvgDBHdem$PercentWhite,AvgDBHdem$AvgDBH)

cor.test(AvgDBHdem$PercentLat_Hisp,AvgDBHdem$AvgDBH)
plot(AvgDBHdem$PercentLat_Hisp,AvgDBHdem$AvgDBH)

cor.test(AvgDBHdem$Population,AvgDBHdem$AvgDBH)
plot(AvgDBHdem$Population,AvgDBHdem$AvgDBH)
```

#coefficient of variation
```{r}
#coefficient of variation for all DBH measurements
NBdbhCV<- data %>% 
  group_by(Neighborhood) %>% 
  summarise(mean = mean(DBH), 
            sd = sd(DBH)) %>% 
  mutate(cv = sd/mean) %>% 
  left_join(NBdem)

```
```{r}
cor.test(NBdbhCV$PercentBlk,NBdbhCV$cv)
plot(NBdbhCV$PercentBlk,NBdbhCV$cv)


cor.test(NBdbhCV$PercentWhite,NBdbhCV$cv)
plot(NBdbhCV$PercentWhite,NBdbhCV$cv)

cor.test(NBdbhCV$PercentLat_Hisp,NBdbhCV$cv)
plot(NBdbhCV$PercentLat_Hisp,NBdbhCV$cv)

cor.test(NBdbhCV$Population,NBdbhCV$cv)
plot(NBdbhCV$Population,NBdbhCV$cv)
```

#Abundance, Richness, evenness for small trees
```{r}
NBabundSM <-data %>% 
  filter(DBH>=0&DBH<=5) %>% 
  group_by(Neighborhood, Species) %>% 
  summarize(abund=sum(present)) 


#calculate richness and eveness for small trees
richevenSM<-community_structure(NBabundSM, abundance.var="abund", time.var= NULL, replicate.var = "Neighborhood") %>%
  left_join(NBdem) %>% 
  unique()
```

#Correlations for small trees
```{r}

#percent white vs richness
cor.test(richevenSM$PercentWhite, richevenSM$richness)
plot(richevenSM$PercentWhite, richevenSM$richness)

#percent black
cor.test(richevenSM$PercentBlk, richevenSM$richness)
plot(richevenSM$PercentBlk, richevenSM$richness)

#percent latino richness
cor.test(richevenSM$PercentLat_Hisp, richevenSM$richness)
plot(richevenSM$PercentLat_Hisp, richevenSM$richness)

#percent white vs evenness
cor.test(richevenSM$Evar,richevenSM$PercentWhite)
plot(richevenSM$PercentWhite, richevenSM$Evar)


#percent black
cor.test(richevenSM$PercentBlk, richevenSM$Evar)
plot(richevenSM$PercentBlk, richevenSM$Evar)


#percent hispanic
cor.test(richevenSM$PercentLat_Hisp, richevenSM$Evar)
plot(richevenSM$PercentLat_Hisp, richevenSM$Evar)
```
#Abundance, Richness, evenness for medium trees
```{r}
NBabundM <-data %>% 
  filter(DBH>5&DBH<20) %>% 
  group_by(Neighborhood, Species) %>% 
  summarize(abund=sum(present)) 


#calculate richness and eveness for medium trees
richevenM<-community_structure(NBabundM, abundance.var="abund", time.var= NULL, replicate.var = "Neighborhood") %>%
  left_join(NBdem) %>% 
  unique()
```

#correlations for medium trees
```{r}
#percent white vs richness
cor.test(richevenM$PercentWhite, richevenM$richness)
plot(richevenM$PercentWhite, richevenM$richness)

#percent black richness
cor.test(richevenM$PercentBlk, richevenM$richness)
plot(richevenM$PercentBlk, richevenM$richness)

#percent white vs evenness
cor.test(richevenM$Evar,richevenM$PercentWhite)
plot(richevenM$PercentWhite, richevenM$Evar)


```

#Abundance, Richness, evenness for large trees
```{r}
NBabundL <-data %>% 
  filter(DBH<=20) %>% 
  group_by(Neighborhood, Species) %>% 
  summarize(abund=sum(present)) 


#calculate richness and eveness for medium trees
richevenL<-community_structure(NBabundL, abundance.var="abund", time.var= NULL, replicate.var = "Neighborhood") %>%
  left_join(NBdem) %>% 
  unique()
```

#correlations for large trees
```{r}
#percent white vs richness
cor.test(richevenL$PercentWhite, richevenL$richness)
plot(richevenL$PercentWhite, richevenL$richness)

#percent white vs evenness
cor.test(richevenL$Evar,richevenL$PercentWhite)
plot(richevenL$PercentWhite, richevenL$Evar)

#percent black
cor.test(richevenL$PercentBlk, richevenL$richness)
plot(richevenL$PercentBlk, richevenL$richness)
```

# neighborhood land coverage csv
```{r}
NBland <-
  # read.csv('../../BaltimoreStreetTreeProject_Large_Data\\neighs.csv') %>%
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\neighs_bnia.csv'
           , col_names = c('id' # call 'community'?
                           , 'n'
                           , 'lc_class')) %>%
  # rename(Number = X357459) %>% 
  filter(lc_class != 0) %>% 
  filter(!is.na(lc_class)) %>% 
  pivot_wider(names_from = lc_class, values_from = n, values_fill = 0) %>% 
  mutate(total_area_in_pixels = rowSums(.[,-1]))
  left_join(NBdem)
head(NBland)
```

# land coverage correlations
```{r}
#Percent white vs land structures
cor.test(NBland$PercentWhite, NBland$Herbaceous)
cor.test(NBland$PercentWhite, NBland$Barren)
cor.test(NBland$PercentWhite, NBland$Structures)
cor.test(NBland$PercentWhite, NBland$`Other Impervious`)
cor.test(NBland$PercentWhite, NBland$Roads)
cor.test(NBland$PercentWhite, NBland$`Tree Canopy over Structures`)
cor.test(NBland$PercentWhite, NBland$`Tree Canopy over Other Impervious`)
cor.test(NBland$PercentWhite, NBland$`Tree Canopy over Roads`)
cor.test(NBland$PercentWhite, NBland$`Tree Canopy`)
cor.test(NBland$PercentWhite, NBland$`Water`)
cor.test(NBland$PercentWhite, NBland$`Tree Canopy`)
cor.test(NBland$PercentWhite, NBland$`Emergent Wetlands`)

#Percent Black vs land structures
cor.test(NBland$PercentBlk, NBland$Herbaceous)
cor.test(NBland$PercentBlk, NBland$Barren)
cor.test(NBland$PercentBlk, NBland$Structures)
cor.test(NBland$PercentBlk, NBland$`Other Impervious`)
cor.test(NBland$PercentBlk, NBland$Roads)
cor.test(NBland$PercentBlk, NBland$`Tree Canopy over Structures`)
cor.test(NBland$PercentBlk, NBland$`Tree Canopy over Other Impervious`)
cor.test(NBland$PercentBlk, NBland$`Tree Canopy over Roads`)
cor.test(NBland$PercentBlk, NBland$`Tree Canopy`)
cor.test(NBland$PercentBlk, NBland$`Water`)
cor.test(NBland$PercentBlk, NBland$`Tree Canopy`)
cor.test(NBland$PercentBlk, NBland$`Emergent Wetlands`)

#Percent Population vs land structures
cor.test(NBland$Population, NBland$Herbaceous)
cor.test(NBland$Population, NBland$Barren)
cor.test(NBland$Population, NBland$Structures)
cor.test(NBland$Population, NBland$`Other Impervious`)
cor.test(NBland$Population, NBland$Roads)
cor.test(NBland$Population, NBland$`Tree Canopy over Structures`)
cor.test(NBland$Population, NBland$`Tree Canopy over Other Impervious`)
cor.test(NBland$Population, NBland$`Tree Canopy over Roads`)
cor.test(NBland$Population, NBland$`Tree Canopy`)
cor.test(NBland$Population, NBland$`Water`)
cor.test(NBland$Population, NBland$`Tree Canopy`)
cor.test(NBland$Population, NBland$`Emergent Wetlands`)

#Percent Latino vs land structures
cor.test(NBland$PercentLat_Hisp, NBland$Herbaceous)
plot(NBland$PercentLat_Hisp, NBland$Herbaceous)
cor.test(NBland$PercentLat_Hisp, NBland$Barren) # r = 0.41
plot(NBland$PercentLat_Hisp, NBland$Barren)
cor.test(NBland$PercentLat_Hisp, NBland$Structures)
plot(NBland$PercentLat_Hisp, NBland$Structures)
cor.test(NBland$PercentLat_Hisp, NBland$`Other Impervious`) # r = 0.4
cor.test(NBland$PercentLat_Hisp, NBland$Roads)
cor.test(NBland$PercentLat_Hisp, NBland$`Tree Canopy over Structures`) 
cor.test(NBland$PercentLat_Hisp, NBland$`Tree Canopy over Other Impervious`)
cor.test(NBland$PercentLat_Hisp, NBland$`Tree Canopy over Roads`)
cor.test(NBland$PercentLat_Hisp, NBland$`Tree Canopy`)
cor.test(NBland$PercentLat_Hisp, NBland$`Water`)
cor.test(NBland$PercentLat_Hisp, NBland$`Tree Canopy`)
cor.test(NBland$PercentLat_Hisp, NBland$`Emergent Wetlands`)

```

#data frame with total area in pixels
```{r}
# (test <- 
#     read_csv('C:\\Users\\trini\\Documents\\Street Trees\\BaltimoreTrees\\BaltimoreStreetTreeProject_Large_Data/neighs.csv'
#              , col_names = c('id', 'n', 'lc_class')) %>%  
#     tidylog::filter(lc_class != 0) %>% 
#     tidylog::filter(!is.na(lc_class)) %>%  
#     tidylog::pivot_wider(names_from = lc_class, values_from = n, values_fill = 0) %>% # need old pipe
#     tidylog::mutate(total_area_in_pixels = rowSums(.[,-1]))                           # for dot notation to work here
#   )
# 
# NBarea <- test %>% 
#   rename(Area = total_area_in_pixels) %>%
#   rename(Neighborhood = id) %>% 
#   left_join(NBdem) %>% 
#   left_join(richeven) %>% 
#   mutate(PopDensity = Population/Area)
```
# area correlations
```{r}
#area and land cover types
cor.test(NBarea$Area, NBarea$`Tree Canopy`)
plot(NBarea$Area, NBarea$`Tree Canopy`)
cor.test(NBarea$Area, NBarea$Herbaceous)
cor.test(NBarea$Area, NBarea$Barren)
cor.test(NBarea$Area, NBarea$Structures)
cor.test(NBarea$Area, NBarea$`Other Impervious`)
cor.test(NBarea$Area, NBarea$Roads)
cor.test(NBarea$Area, NBarea$`Tree Canopy over Structures`)
cor.test(NBarea$Area, NBarea$`Tree Canopy over Other Impervious`)
cor.test(NBarea$Area, NBarea$`Tree Canopy over Roads`)
cor.test(NBarea$Area, NBarea$Water)

#roads and land cover
cor.test(NBarea$Roads, NBarea$`Tree Canopy`)
plot(NBarea$Roads, NBarea$`Tree Canopy`)
cor.test(NBarea$Roads, NBarea$Herbaceous)
cor.test(NBarea$Roads, NBarea$Barren)
cor.test(NBarea$Roads, NBarea$Structures)
cor.test(NBarea$Roads, NBarea$`Other Impervious`)
cor.test(NBarea$Roads, NBarea$`Tree Canopy over Structures`)
cor.test(NBarea$Roads, NBarea$`Tree Canopy over Other Impervious`)
cor.test(NBarea$Roads, NBarea$`Tree Canopy over Roads`)

#roads and richness/evenness
cor.test(NBarea$Roads, NBarea$richness)
cor.test(NBarea$Roads, NBarea$Evar)

#density vs richness/evenness
cor.test(NBarea$PopDensity, NBarea$richness)
cor.test(NBarea$PopDensity, NBarea$Evar)
```
#calculate land cover percentages
```{r}

```

