---
title: "Street Trees Independent Variables"
author: "Trini et al"
date: "`r format(Sys.Date())`"
output: html_document
---
This file contains independent variables for street tree data anaylsis 
Uses BNIA neighborhood delineations or Community Statistical Area (CSA)

# 0 set up
```{r include=FALSE}
# load packages

library(tidyverse)
library(sf) #simple features - vector GIS support
library(mapview) #quick web maps 
library(tidylog) #makes dplyr and tidyr verbose
library(dplyr)

#working directory
getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)
```


#1 read in BNIA csv files with demographic data

#BNIA CSA race demographics ## use 2020
```{r include=FALSE}

PercentAsian <-  
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Asian_(Non-Hispanic).csv') %>%   select(OBJECTID, CSA2010, pasi20) %>% 
  rename(PercAsian = pasi20)

PercentLatino <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Hispanic.csv') %>% 
  select(OBJECTID, CSA2010, phisp20) %>% 
  rename(PercLatino = phisp20)

PercentBlk <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_Black_African-American_(Non-Hispanic).csv') %>% 
  select(OBJECTID, CSA2010, paa20) %>% 
  rename(PercBlk = paa20)

PercentWhite <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Residents_-_White_Caucasian_(Non-Hispanic).csv') %>% 
  select(OBJECTID, CSA2010, pwhite20) %>% 
  rename(PercWhite = pwhite20) 

```

#BNIA total population by CSA
```{r}
TotalPop <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Total_Population.csv') %>% 
  select(OBJECTID, CSA2010, tpop20)
```

#BNIA percent residential properties that are vacant by CSA
```{r}
Vacant <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percentage_of_Residential_Properties_that_are_Vacant_and_Abandoned.csv') %>% 
  select(OBJECTID, CSA2010, vacant20)
```

#BNIA median income
```{r}
Income <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Median_Household_Income.csv') %>% 
  select(OBJECTID, CSA2010, mhhi20) 
```

# BNIA percent below poverty line
```{r}
Poverty <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_of_Family_Households_Living_Below_the_Poverty_Line.csv') %>% 
  select(OBJECTID, CSA2010, hhpov21) 

```

# percent of population 25 and over with a Bachelor's Degree or above
```{r}
Edu <- 
  read_csv('../../BaltimoreStreetTreeProject_Large_Data\\Percent_Population_(25_Years_and_over)_with_a_Bachelor_s_Degree_or_Above.csv') %>% 
  select(OBJECTID, CSA2010, bahigher19) 

```

# Land Cover BNIA total NB area
```{r}
LandCover <- read_csv('../../BaltimoreStreetTreeProject_Large_Data\\neighs_bnia.csv'
           , col_names = c('CSA2010'
                           , 'n'
                           , 'lc_class')) %>%
  filter(lc_class != 0) %>% 
  filter(!is.na(lc_class)) %>% 
  pivot_wider(names_from = lc_class, values_from = n, values_fill = 0) %>% 
  mutate(total_area_in_pixels = rowSums(.[,-1]))
```


# Population density
```{r}
PopDensity <- LandCover %>% 
  left_join(TotalPop) %>% 
  rename(Pop = tpop20) %>% 
  mutate(PopDensity = Pop/total_area_in_pixels) %>% 
  select(CSA2010, PopDensity)
```



# % impervious land cover
```{r}
PercentLand <- LandCover %>% 
  mutate(Imperv = Structures + `Other Impervious`+ Roads + `Tree Canopy over Structures` + `Tree Canopy over Other Impervious` + `Tree Canopy over Roads` ) %>% 
  mutate(PercentImp = (Imperv/total_area_in_pixels)*100) %>% 
  mutate(Pervious = Water + `Tree Canopy`+ `Scrub\\Shrub`  
         + Herbaceous + Barren + `Emergent Wetlands`) %>% 
  mutate(PercPervious = (Pervious/total_area_in_pixels)*100)
```


# % canopy tree cover
```{r}
PercentCanopy <- LandCover %>% 
  mutate(PercCanopy = (`Tree Canopy`/total_area_in_pixels)*100)
```

# total road length (not including highways)
```{r}

bmore_bnia <- st_read('../../BaltimoreStreetTreeProject_Large_Data\\Community Statistical Areas (BNIA neighborhoods)/Community_Statistical_Areas_(CSAs)__Reference_Boundaries.shp') |> #BNIA community delineations
  st_make_valid()

roads <- 
  st_read('../../BaltimoreStreetTreeProject_Large_Data\\Baltimore Roads/MDOT_Know_Your_Roads.shp'
, as_tibble = TRUE) %>% 
   st_make_valid() |> 
   st_transform(crs = st_crs(bmore_bnia))

NBRoads <- bmore_bnia %>%
  select(CSA2020, Community) %>% 
  st_intersection(roads %>% mutate(road_length_ft = as.double(st_length(.)))) %>% 
  st_drop_geometry() %>% 
  group_by(CSA2020) %>% 
  summarize(sum_road_length_ft = sum(road_length_ft)) %>% 
  rename(CSA2010 = CSA2020)

NBRoads %>% 
  ggplot(aes(sum_road_length_ft) ) +
  geom_histogram()

NBRoads %>% 
  arrange(desc(sum_road_length_ft))
NBRoads %>% 
  arrange(sum_road_length_ft)

```

# temperature
```{r}


```

# percent immigrants (NOT in BNIA - in the census) 

# 2 join all independent variables 
```{r}
IndVar <- PercentAsian %>% 
  left_join(PercentLatino, by = "CSA2010") %>% 
  left_join(PercentBlk, by = "CSA2010") %>%
  left_join(PercentWhite, by = "CSA2010") %>% 
  left_join(TotalPop, by = "CSA2010") %>% 
  left_join(Income, by = "CSA2010") %>%
  left_join(Poverty, by = "CSA2010") %>% 
  left_join(Edu, by = "CSA2010") %>% 
  left_join(PopDensity) %>% 
  left_join(PercentCanopy, by = "CSA2010") %>% 
  left_join(PercentLand, by = "CSA2010") %>% 
  left_join(NBRoads, by = "CSA2010")
  
  
```



