---
title: "New Analysis"
output:
  pdf_document: default
  html_document:
    fig_width: 8
    fig_height: 5
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
---
Scatter plots for correlations p<0.05

<!-- set up -->
```{r include=FALSE}
# load packages
library(tidyverse)
library(sf) 
library(mapview)  
library(tidylog) 
library(purrr)
library('correlation')
library("writexl")

#working directory
getwd()
list.files()
list.files('../../BaltimoreStreetTreeProject_Large_Data/')
list.files('../../BaltimoreStreetTreeProject_Large_Data/', recursive = TRUE)

var_names <- as_labeller(
     c(`PercBlk` = "% Black", `PercWhite` = "% White",`mhhi20` = "Income", 
       `bahigher19` = "Bachelor's or higher", `avg_temp` = "Average Temperature"))

var_names_2 <- as_labeller(
     c(`AvgDBH` = "Average DBH", `Evar` = "Evenness",`Perc32` = "Ideal Pits", 
       `PercNat` = "Nativity", `PercGood` = "% Good", `PercPoor` = "% Poor", `percent_filled` = "% Filled", `richness` = "Richness"))
```

#1 Combined dep+ind
```{r include=FALSE}
Dependent <- read_csv('../input_data/Dependent_variable.csv')
Independent<- read_csv('../input_data/Independent_variable.csv')

Combined <- Dependent %>% 
  left_join(Independent, by = "CSA2020") 

cor.test(Combined$avg_temp, Combined$spave)
```

#2 All p values and correlations
```{r include=FALSE}
 
#empty df to store results
cor_results <- data.frame()

# Get names of all variables in df

Vars <- Combined %>% 
  select(where(is.numeric)) 

variables <- names(Vars)

# Loop through all combinations of variables
for (i in 1:(length(variables) - 1)) {
  for (j in (i + 1):length(variables)) {
    
    # Extract the two variables for the current combination
    var1 <- Vars[[variables[i]]]
    var2 <- Vars[[variables[j]]]
    
    # Run correlation test
    cor_test_result <- cor.test(var1, var2)
    
    # Extract correlation coefficient and p-value
    correlation_coefficient <- cor_test_result$estimate
    p_value <- cor_test_result$p.value
    
    # Create a row for the results dataframe
    result_row <- data.frame(
      Variable1 = variables[i],
      Variable2 = variables[j],
      CorrelationCoefficient = correlation_coefficient,
      PValue = p_value
    )
    
    # Bind row to results df
    cor_results <- rbind(cor_results, result_row)
  }
}

# View the resulting dataframe
print(cor_results)

# Different Method
# # like this
# correlation::correlation(Vars)
# 
# # # or like this,
# pvalues <- 
# Combined |> 
#   select(where(is.numeric)) |> 
#   correlation::correlation() |>
#   tibble() # can add this 
```

#3 Filtered p < 0.05
```{r message=FALSE, include=FALSE}

p_filtered_0.05 <- cor_results %>% 
  filter(PValue<0.05) 

```

<!-- ## Filtered p < 0.001 -->
```{r message=FALSE, warning=FALSE, include=FALSE}
p_filtered_0.001 <- cor_results %>% 
  filter(PValue<0.001)

# knitr:: kable(p_filtered_0.001)
```

##Total Tree Pits
```{r}
totpits_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "all_sites" | Variable2 == "all_sites")

view(totpits_r)
```

## Stocking
```{r echo=FALSE, message=FALSE, warning=FALSE}
stocking_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "percent_filled" | Variable2 == "percent_filled")

view(stocking_r)

perc_filled_plot <- Combined |> 
  tidylog::select(
                   percent_filled  # depend
                  , PercBlk
                  , PercWhite
                  , bahigher20
                  , mhhi20
                  ) |> 
  pivot_longer(-percent_filled) |> 
  ggplot(aes(value, percent_filled)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, nrow =2
             , labeller = var_names
             ) +
  labs(
         x = 'Predictor'
       , y = 'Percent Filled'
       ) +   theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

percfill_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "percent_filled" | Variable2 == "percent_filled")

knitr:: kable(percfill_r)

ggsave("perc_filled_plot.png", perc_filled_plot, dpi = 300)
```

##Richness
```{r}
richtot_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "rich_tot" | Variable2 == "rich_tot")

view(richtot_r)
```

## Evenness
```{r echo=FALSE, message=FALSE, warning=FALSE}
even_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Evar" | Variable2 == "Evar")

view(even_r)

even_plot <- Combined |> 
  tidylog::select(
                    Evar  # depend 
                  , PercBlk
                  , avg_temp
                  , bahigher20
                  ) |> 
  pivot_longer(-Evar) |> 
  ggplot(aes(value, Evar)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Trees Evenness'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

Evar_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Evar" | Variable2 == "Evar")

knitr:: kable(Evar_r)

ggsave("even_plot.png", even_plot, dpi = 300)
```

##Beta Diversity
```{r}
betadiv_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "spave" | Variable2 == "spave")

view(betadiv_r)

```

## Average DBH
```{r echo=FALSE, message=FALSE, warning=FALSE}
dbh_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "AvgDBH" | Variable2 == "AvgDBH")

view(dbh_r)

dbh_plot <- Combined |> 
  tidylog::select(
                    AvgDBH  # depend 
                  , PercBlk
                  , avg_temp
                  ) |> 
  pivot_longer(-AvgDBH) |> 
  ggplot(aes(value, AvgDBH)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Average DBH'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

AvgDBH_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "AvgDBH" | Variable2 == "AvgDBH")

knitr:: kable(AvgDBH_r)

ggsave("dbh_plot.png", dbh_plot, dpi = 300)
```


## Percent Trees in Poor Condition
```{r echo=FALSE, message=FALSE, warning=FALSE}
poor_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercPoor" | Variable2 == "PercPoor")

view(poor_r)

percpoor_plot <- Combined |> 
  tidylog::select(
                    PercPoor  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher20
                  ) |> 
  pivot_longer(-PercPoor) |> 
  ggplot(aes(value, PercPoor)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Trees in Poor Condition'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

percpoor_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercPoor" | Variable2 == "PercPoor")

knitr:: kable(percpoor_r)

ggsave("percpoor_plot.png", percpoor_plot, dpi = 300)
```

## Percent Trees in Good Condition
```{r echo=FALSE, message=FALSE, warning=FALSE}
good_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercGood" | Variable2 == "PercGood")

view(good_r)

percgood_plot <- Combined |> 
  tidylog::select(
                    PercGood  # depend 
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  , bahigher19
                  ) |> 
  pivot_longer(-PercGood) |> 
  ggplot(aes(value, PercGood)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Trees in Good Condition'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

percgood_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercGood" | Variable2 == "PercGood")

knitr:: kable(percgood_r)

ggsave("percgood_plot.png", percgood_plot, dpi = 300)
```

##  Percent Tree Pits > 32 sq ft
```{r echo=FALSE, message=FALSE, warning=FALSE}
perc32_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Perc32" | Variable2 == "Perc32")

view(perc32_r)

plot_32 <- Combined |> 
  tidylog::select(
                    Perc32  # depend 
                  , bahigher20
                  ) |> 
  pivot_longer(-Perc32) |> 
  ggplot(aes(value, Perc32)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, labeller = var_names
             ) +
  labs(
        x = 'Predictor'
       , y = 'Percent Tree Pits >32 sq. ft.'
       ) + 
  
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL

perc32_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "Perc32" | Variable2 == "Perc32")

knitr:: kable(perc32_r)

ggsave("avgtemp_plot.png", temp_plot, dpi = 300)
```

## Temperature 
```{r}
temp_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "avg_temp" | Variable2 == "avg_temp")
view(temp_r)

temp_plot <- Combined |> 
  tidylog::select(
    avg_temp# depend 
    , Evar
    , AvgDBH
    , PercNat
    , Perc32
  ) |> 
  pivot_longer(-avg_temp) |> 
  ggplot(aes(x = avg_temp, y = value)) +  # Swap x and y aesthetics
  geom_point() +
  geom_smooth(method = 'lm') +  # forces a linear model
  facet_wrap(~name, scales = 'free_y', ncol = 2, labeller = var_names_2) +  # scales = 'free_y'
  labs(
    x = 'Average Temperature',  # Updated x-axis label
    y = 'Response'  # Updated y-axis label
  ) + 
  theme_bw() +
  theme(panel.grid = element_blank(), text = element_text(size = 20)) +
  NULL

ggsave("avgtemp_plot.png", temp_plot, dpi = 300)
```

## Flowering 
```{r}
percflor_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercFlor" | Variable2 == "PercFlor")

view(percflor_r)

Combined |> 
  tidylog::select(
                   PercShowy  # depend
                  , PercBlk
                  , PercWhite
                  , mhhi20
                  ) |> 
  pivot_longer(-PercShowy) |> 
  ggplot(aes(value, PercShowy)) + 
  geom_point() +
  geom_smooth(method = 'lm') + # forces a linear model
  facet_wrap(~name, scales = 'free_x'
             , ncol = 2, nrow =2
             , labeller = var_names
             ) +
  labs(
         x = 'Predictor'
       , y = 'PercShowy'
       ) +   
  theme_bw() +
  theme(panel.grid=element_blank(),
        text=element_text(size=20)) +
  NULL
```

## Percent Showy
```{r}
percshow_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercShowy" | Variable2 == "PercShowy")

view(percshow_r)
```

## Average Showiness
```{r}
avgshow_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "AvgShowy" | Variable2 == "AvgShowy")

view(avgshow_r)
```

## color richness
```{r}
color_rich_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "color_rich" | Variable2 == "color_rich")

view(color_rich_r)
```

# r value matrix
```{r}
cor_matrix <- cor_results %>% 
  select(Variable1, Variable2, CorrelationCoefficient) %>% 
  spread(Variable2, CorrelationCoefficient) %>% 
  column_to_rownames(var = "Variable1")

print(cor_matrix)


write_xlsx(cor_matrix, '..\\..\\BaltimoreStreetTreeProject_Large_Data\\Rmatrix.xlsx')

```

#income
```{r}
income_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "mhhi20" | Variable2 == "mhhi20")

view(income_r)
```

#education
```{r}
education_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "bahigher20" | Variable2 == "bahigher20")
```

# Race
```{r}
pblack_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercBlk" | Variable2 == "PercBlk")
view(pblack_r)

pwhite_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "PercWhite" | Variable2 == "PercWhite")
view(pwhite_r)
```

# Vacancy
```{r}
vacancy_r <- p_filtered_0.05 %>% 
  filter(Variable1 == "vacant20" | Variable2 == "vacant20")
view(vacancy_r)


vacancy_plot <- Combined |> 
  tidylog::select(
    vacant20# depend 
    , richness
    , AvgDBH
    , PercGood
    , Perc32
  ) |> 
  pivot_longer(-vacant20) |> 
  ggplot(aes(x = vacant20, y = value)) +  # Swap x and y aesthetics
  geom_point() +
  geom_smooth(method = 'lm') +  # forces a linear model
  facet_wrap(~name, scales = 'free_y', ncol = 2, labeller = var_names_2) +  # scales = 'free_y'
  labs(
    x = 'Vacancy',  # Updated x-axis label
    y = 'Response'  # Updated y-axis label
  ) + 
  theme_bw() +
  theme(panel.grid = element_blank(), text = element_text(size = 20)) +
  NULL

ggsave("vacancy_plot.png", vacancy_plot, dpi = 300)
```

